---
title: "01_data_processing.Rmd"
author: "Sammy Stone"
date: "2024-06-12"
output: html_document
embed-resources: true
---
# RADIOHEAD Data Processing
#
# This script processes medication and clinical data to create analysis-ready dataframes:
# - Merges harmonized medications with primary RADIOHEAD dataframe
# - Filters to systemic medications
# - Calculates polypharmacy 
# - Assigns medication subclassses
#
# Key Outputs:
# - RADIOHEAD_multivariate_df.csv (main analysis dataframe, n=1024)
# - RADIOHEAD_systemicdrugs.csv
# - RADIOHEAD_polypharmacy_df.csv
# - RADIOHEAD_subclasses.csv
# - individual_drugs_df.csv
# - RADIOHEAD_demographic_df.csv

# ==============================================================================
# Setup
# ==============================================================================
## Load packages
```{r}
library(here)
library(tidyverse)
library(readxl)
library(ggplot2)
library(stringr)
library(dplyr)
library(readr)
library(readxl)
library(openxlsx)
```
# Load project configuration
```{r}
source(here::here("R", "config.R"))
```
## Utility functions
```{r}
theme_custom <- 
  theme_bw() +
  theme(panel.grid = element_blank())

# Make first letter capital
simpleCap <- function(x) {
  x <- tolower(x)
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
        sep="", collapse=" ")
}
# https://stackoverflow.com/questions/2261079/how-can-i-trim-leading-and-trailing-white-space
# Returns string without leading or trailing white space
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```


# ==============================================================================
# Load Raw Data
# ==============================================================================

```{r load-data}
# Harmonized medications (from ZW)
man_h7 <- read_excel(here::here("data", "raw", "Per_patient_harmonized_round_10.xlsx"))

# Clinical metadata
dat_meta <- read.csv(here::here("data", "raw", "df.clean15.postMS_2024-08-22.csv"))

message(sprintf("✓ Loaded %d medication records", nrow(man_h7)))
message(sprintf("✓ Loaded %d patient records", nrow(dat_meta)))
```
# Merged_df-From ZW

generate "merged_df" a dataframe that combines meds dataframe (per patient harmonized) with clinical information (df clean)

```{r}
head(man_h7)
head(dat_meta)
#Prepare medication data
df1 <- man_h7 %>%
  select(record_id, Revised_Classification) %>%
  mutate(Revised_Classification = gsub("-", "_", Revised_Classification),
         Revised_Classification = gsub(" ", "_", Revised_Classification))
  View(dat_meta)
```


```{r}
#Select clinical data
df2 <- dat_meta %>%
select(subject.id, meddra.disease.preferred.name, os, subject.dead.hospice, subject.age, line.1.therapy, prior.targeted_therapy, subject.metastatic.disease, irae.any, irae.severe, stage)
  
# Merge data frames according to patient id
merged_df1 <- left_join(df1, df2, by = c("record_id" = "subject.id"))

head(merged_df1)
```

```{r}
# Create binary columns for each unique value in Class_1
unique_Revised_Classification <- unique(df1$Revised_Classification)
```

```{r}
# Integrate dataframes
merged_df1 <- left_join(df1, df2, by = c("record_id" = "subject.id")) %>%

  mutate(

    Revised_Classification =

      case_when(Revised_Classification == "NA" ~ NA,

    TRUE ~ Revised_Classification))
head(merged_df1)
```


```{r}
merged_df <- merged_df1 %>%
  distinct() %>%
  mutate(Revised_Classification_present = 1) %>%
  pivot_wider(names_from = Revised_Classification, values_from = Revised_Classification_present, values_fill = 0)

colnames(merged_df)
```

```{r}
#Exclude NA values
merged_df <- merged_df %>%
  filter(stage != "NA")
```

```{r}
message(sprintf("✓ Created dataframe with %d patients and %d drug classes", 
                nrow(merged_df), ncol(merged_df) - 11))

# Save merged dataframe
write.csv(
  merged_df,
  here::here("data", "processed", "RADIOHEAD_1070_meds_df.csv"),
  row.names = FALSE
)

message("✓ Saved RADIOHEAD_1070_meds_df.csv") 

#1 patient has been removed by collaborator in data cleanup, now 1069 pts instead of 1070
```

## Merged df2
Merged df but with individual Drugs
```{r}
head(man_h7)
head(dat_meta)
df1_2 <- man_h7 %>%
  select(record_id, harmonized_final) %>%
  mutate(harmonized_final = gsub("-", "_", harmonized_final),
         harmonized_final = gsub(" ", "_", harmonized_final))
#df2 from above
```

```{r}
# Create binary columns for each unique value in harmonized final
unique_harmonized_final <- unique(df1_2$harmonized_final)

# Merge data frames based on record_id and subject.id

merged_df2 <- left_join(df1_2, df2, by = c("record_id" = "subject.id")) %>%

  mutate(

    harmonized_final =

      case_when(harmonized_final == "NA" ~ NA,

    TRUE ~ harmonized_final))

head(merged_df2) 

merged_df2 <- merged_df2 %>%
  distinct() %>%
  mutate(harmonized_final_present = 1) %>%
  pivot_wider(names_from = harmonized_final, values_from = harmonized_final_present, values_fill = 0)
colnames(merged_df2) 

```
individual drugs 
```{r}
merged2_filt <- merged_df2 %>%
    filter(!(stage == "UNK"))%>%
 filter(!is.na(subject.dead.hospice)) 

merged2_stage <- merged2_filt %>%
    mutate(stage2 = case_when(
    stage %in% c("0","i", "ii") ~"0-ii",
    TRUE ~ stage)) 

merged2_stage2 <- merged2_stage %>%
  select(-c("stage", "irae.any", "irae.severe"))

individual_with_stage <- merged2_stage2 %>%
  rename(stage = stage2)
```


```{r}
merged2_update <- merged2_stage2 %>%
  rename(stage = stage2)
```

```{r}
individual_drugs <- merged2_update %>%
mutate(Treatment = case_when(
    line.1.therapy == "chemo+pd1-pdl1" ~ "chemo+ICI",
    line.1.therapy == "chemo+ctla4+pd1-pdl1" ~ "chemo+ICI",
    TRUE ~ line.1.therapy))
```

```{r}
individual_drugs <- individual_drugs %>%
  mutate(cancer_agreggate = ifelse(meddra.disease.preferred.name %in% 
c("Non-small cell lung cancer", "Malignant melanoma", "Renal cell carcinoma", "Bladder cancer", "Small cell lung cancer"),
meddra.disease.preferred.name,
"Other"))
```

```{r}
# Save individual drugs dataframe
write.csv(
  individual_drugs,
  here::here("data", "processed", "individual_drugs_df.csv"),
  row.names = FALSE
)
```

# ==============================================================================
# Create Cohort and Systemic Medications Dataframe
# ==============================================================================
Cohort df = Merged_df patients with no missing death event (subject.dead.hospice)
1028 patients in newest df update
```{r}
# Filter to patients with known death events
cohort <- merged_df %>%
  filter(!is.na(subject.dead.hospice))
nrow(cohort)
```

Corticosteroid DF
Create corticosteroid column with corticosteroids unclear + corticosteroids systemic combined
```{r}
# Combine corticosteroid columns
merged_w_cort <- cohort %>%
  mutate(Corticosteroids = if_else(Corticosteroid_systemic == 1 | Corticosteroid_unclear == 1, 1, 0))

#View(merged_w_cort)
```

#Systemic DF
filter out drug classes we're not interested in 
```{r}
#filter out drug classes we're not interested in--no vitamins/supplements or non systemic drugs

systemic_df <-  merged_w_cort %>%
         select(-c("Vitamin", "NA", "Drop_NA", "Supplement", "Cancer", "Anesthetic_non_systemic", "Anti_glaucoma_non_systemic", "Cholinergic_non_systemic", "Mast_cell_stabilizer_non_systemic", "Supplement_non_systemic", "Antibiotics_non_systemic", "Corticosteroid_inhaled", "Anti_histamine_non_systemic", "Decongestant_non_systemic", "Anti_histamine_non_systemic", "Corticosteroid_non_systemic", "Anti_fungal_non_systemic", "Bronchodilator", "Cannabinoid"))
```

```{r}
# Save systemic drugs dataframe
write.csv(
  systemic_df,
  here::here("data", "processed", "RADIOHEAD_systemicdrugs.csv"),
  row.names = FALSE
)

message(sprintf("✓ Saved RADIOHEAD_systemicdrugs.csv with %d patients", nrow(systemic_df)))
```
# ==============================================================================
# Calculate Polypharmacy
# ==============================================================================

#Calculate polypharmacy
```{r}
# Prepare for polypharmacy calculation-- exclude duplicate medications
polypharmacydf <- systemic_df %>%
  select(-c("Corticosteroid_unclear", "Corticosteroids"))

# Save these columns for later
excluded_df <- systemic_df %>%
  select(record_id, Corticosteroid_unclear, Corticosteroids)

#Calculate polypharmacy
polydf <- polypharmacydf %>%
  mutate(
    polypharmacy = rowSums(polypharmacydf[12:81]),
    polypharmacy_binary = polypharmacy >= 5
  )
```

```{r}
# Save list of drug classes used in polypharmacy calculation 
polypharmacy_drugclasses <- as.data.frame(names(polypharmacydf[12:81]))
colnames(polypharmacy_drugclasses) <- "DrugClass"

write.xlsx(
  polypharmacy_drugclasses,
  here::here("output", "tables", "Drugs_Used_in_Polypharmacy_Calculation.xlsx")
)

message("✓ Saved Drugs_Used_in_Polypharmacy_Calculation.xlsx (polypharmacy drug classes)")
```

```{r}
# Save polypharmacy dataframe
write.csv(
  polydf,
  here::here("data", "processed", "RADIOHEAD_polypharmacy_df.csv"),
  row.names = FALSE
)

message(sprintf("✓ Saved RADIOHEAD_polypharmacy_df.csv"))
message(sprintf("  Mean polypharmacy: %.2f", mean(polydf$polypharmacy)))
message(sprintf("  Median polypharmacy: %.0f", median(polydf$polypharmacy)))
```

# ==============================================================================
# Create Subclass Columns
# ==============================================================================

Using merged_df2, we'll identify relevant sub-classes from individual medications. Then we'll merge with merged_df

#Assign classes
```{r}
#mental health
benzodiazepines <- c("Clonazepam", "Alprazolam", "Lorazepam", "Diazepam", "Temazepam", "Chlordiazepoxide", "Clobazam")

ssris <- c("Escitalopram", "Citalopram", "Fluoxetine", "Paroxetine", "Sertraline")

tcas <- c("Amitriptyline", "Nortriptyline", "Doxepin") 

atypicals <- c("Bupropion", "Mirtazapine", "Buspirone")
  
snris <- c("Duloxetine", "Venlafaxine", "Desvenlafexine")

#cardiovascular

beta_blockers <- c("Metoprolol", "Carvedilol", "Atenolol", "Propranolol", "Sotalol", 
                   "Nebivolol", "Nadolol", "Labetalol", "Bisoprolol")

anti_arrythmics <- c("Amiodarone", "Flecainide", "Dofetilide", 
                      "Mexiletine", "Propafenone")

ACEi_ARB <- c("Lisinopril", "Enalapril", "Benazepril", "Ramipril", "Quinapril",
          "Fosinopril", "Perindopril", "Trandolapril", "Losartan", "Valsartan", "Irbesartan", "Olmesartan", "Candesartan", "Azilsartan","Telmisartan")

cc_blockers <- c("Amlodipine", "Diltiazem", "Nifedipine", "Felodipine", "Verapamil")
  
  
#Cholesterol Lowing
statins <- c("Atorvastatin","Simvastatin", "Rosuvastatin", "Pravastatin", "Lovastatin", "Fluvastatin")

fibrates <- c("Gemfibrozil", "Fenofibrate")

#Diabetes

dppi_4i <- c("Sitagliptin", "Linagliptin", "Alogliptin", "Saxagliptin") 

sulfonylurea <- c("Glipizide", "Glimepiride", "Glyburide", "Glibenclamide")

glp_1 <- c("Liraglutide", "Dulaglutide", "Exenatide", "Semaglutide")

#Anti-pyretic

nsaid <- c("NSAID", "Ibuprofen", "Naproxen", "Meloxicam", "Diclofenac", "Celecoxib", "Indomethacin", 
           "Ketorolac", "Nabumetone", "Nepafenac")

#bronchodilator 

b2_ag <- c("Vilanterol", "Formoterol", "Arformoterol", "Olodaterol", "Salmeterol", "Albuterol", "Levosalbutamol", "Levalbuterol")
  
anti_cholinergic <- c("Umeclidinium", "Tiotropium") 

#anti-emetic

sert_atg <- c("Ondansetron", "Granisetron", "Palonosetron")

dopamine_atg <- c("Prochlorperazine", "Promethazine") 

#Anti-coagulants

DOAC <- c("Apixaban", "Rivaroxaban", "Dabigatran")

anti_platelet <- c("Aspirin", "Clopidogrel", "Cilostazol", "Ticagrelor", "Prasugrel", "Dipyridamole")

heparin_LMWH <- c("Enoxaparin", "Heparin", "Dalteparin", "Fondaparinux")
```

Incorporate subclasses in merged_df2
```{r}
subdf <- merged_df2 %>%
  mutate(Benzodiazepines = ifelse(rowSums(select(., all_of(benzodiazepines))) > 0, 1, 0),
        SSRIs = ifelse(rowSums(select(., all_of(ssris))) > 0, 1, 0),
        TCAs = ifelse(rowSums(select(., all_of(tcas))) > 0, 1, 0),
        #Atypicals = ifelse(rowSums(select(., all_of(atypicals))) > 0, 1, 0), 
        #SNRIs = ifelse(rowSums(select(., all_of(snris))) > 0, 1, 0),
        Beta_Blockers = ifelse(rowSums(select(., all_of(beta_blockers))) > 0, 1, 0),
        ACEi_ARB = ifelse(rowSums(select(., all_of(ACEi_ARB))) > 0, 1, 0),
        #Anti_Arrythmics = ifelse(rowSums(select(., all_of(anti_arrythmics))) > 0, 1, 0),
        #Sulfonylurea = ifelse(rowSums(select(., all_of(sulfonylurea))) > 0, 1, 0),
        NSAIDs = ifelse(rowSums(select(., all_of(nsaid))) > 0, 1, 0),
        #B2_ag = ifelse(rowSums(select(., all_of(b2_ag))) > 0, 1, 0),
        #MA_antag = ifelse(rowSums(select(., all_of(anti_cholinergic))) > 0, 1, 0),
        #Sert_atg = ifelse(rowSums(select(., all_of(sert_atg))) > 0, 1, 0),
        #Dopamine_atg = ifelse(rowSums(select(., all_of(dopamine_atg))) > 0, 1, 0),
        #GLP_1 = ifelse(rowSums(select(., all_of(glp_1))) > 0, 1, 0),
        #DPPI_4i = ifelse(rowSums(select(., all_of(dppi_4i))) > 0, 1, 0),
        #Fibrates = ifelse(rowSums(select(., all_of(fibrates))) > 0, 1, 0),
        CC_blockers = ifelse(rowSums(select(., all_of(cc_blockers))) > 0, 1, 0),
        #DOACs = ifelse(rowSums(select(., all_of(DOAC))) > 0, 1, 0),
        #Anti_platelet = ifelse(rowSums(select(., all_of(anti_platelet))) > 0, 1, 0),
        Statin = ifelse(rowSums(select(., all_of(statins))) > 0, 1, 0))
        #Heparin_LMWH = ifelse(rowSums(select(., all_of(heparin_LMWH))) > 0, 1, 0))
        
        
```

remove non subclass columns from subdf

```{r}
##only subclass script below
#indv_drug_col <- names(subdf)[2:558 ]

#subdf <- subdf %>%
  #select(-all_of(indv_drug_col)) 

##including lorazepam and diazepam

subdf <- subdf %>%
  select(-c(2:558))

colnames(subdf)
```

merge subdf with all med classes 

```{r}
cohortsub <- left_join(polydf, subdf, by = "record_id") 
```
remove na, vitamins, and supplements columns for analyses- already done in systemic_df

Sanity check 
```{r}
colnames(cohortsub)
```

```{r}
# Save subclasses dataframe
write.csv(
  cohortsub,
  here::here("data", "processed", "RADIOHEAD_subclasses.csv"),
  row.names = FALSE
)

message("✓ Saved RADIOHEAD_subclasses.csv with subclass columns")
```


# ==============================================================================
# Create Final Multivariate Analysis Dataframe
# ==============================================================================

```{r}
#Combine cancer stages 0-1
cohortsub <- cohortsub %>%
  mutate(stage2 = case_when(
    stage %in% c("0","i", "ii") ~"0-ii",
    TRUE ~ stage))

```

```{r}
#Remove columns not included in multivariate analysis
cohortstage <- cohortsub %>%
  select(-c("stage", "irae.any", "irae.severe", "polypharmacy_binary")) %>%
  rename(stage = stage2)
```

```{r}
#Reorder df to put polypharmacy at the top, add back in columns omitted from polypharmacy calculation, then get drug class counts 
cohort_ordered <- cohortstage %>%
  select(1:8, stage, polypharmacy, 9:88)
```

```{r}
#Join back excluded columns
poly_cohort <- left_join(cohort_ordered, excluded_df, by = "record_id")
```

```{r}
#Restrict df to drug classes with at least 50 patients on them 

colsums <- colSums(poly_cohort[11:90])

sub <- names(colsums[colsums >= 50])

sub_df <- poly_cohort %>%
 select(1:10, all_of(sub))
```


```{r}
#exclude patients where stage = UNK 
sub_df <- sub_df %>% 
  filter(!(stage == "UNK"))

```
*now have 1024 patients for use in multivariate analysis 

```{r}
#Combine treatments
sub_df <- sub_df %>%
  mutate(Treatment = case_when(
    line.1.therapy == "chemo+pd1-pdl1" ~ "chemo+ICI",
    line.1.therapy == "chemo+ctla4+pd1-pdl1" ~ "chemo+ICI",
    TRUE ~ line.1.therapy))

unique(sub_df$Treatment) 
```


```{r}
#Sanity checks
check_df <- sub_df %>%
  select(record_id, Treatment, line.1.therapy)
View(check_df)
```

```{r}
#Remove old line.1.therapy columns
sub_df <- sub_df %>%
  select(-line.1.therapy)
```

```{r}
#Add new cancer column + sub-class column 
sub_df <- sub_df %>%
mutate(cancer_agreggate = ifelse(meddra.disease.preferred.name %in% 
c("Non-small cell lung cancer", "Malignant melanoma", "Renal cell carcinoma", "Bladder cancer", "Small cell lung cancer"),
meddra.disease.preferred.name,
"Other"))
```

reorder columns 
```{r}
sub_df <- sub_df %>%
  select(1:5, Treatment, cancer_agreggate, 6:41)
head(sub_df)
colnames(sub_df)
```
```{r}
# Save main multivariate dataframe
write.csv(
  sub_df,
  here::here("data", "processed", "RADIOHEAD_multivariate_df.csv"),
  row.names = FALSE
)

message(sprintf("✓ Saved RADIOHEAD_multivariate_df.csv with %d patients", nrow(sub_df)))
message(sprintf("  Included %d drug classes with ≥50 patients", ncol(sub_df) - 10))
```

# ==============================================================================
# Create Demographic Dataframe
# ==============================================================================

#Demographic DF

```{r}
#Import additional information from dat_meta 
df3 <- dat_meta %>%
select(subject.id, subject.sex, subject.ethnicity, subject.race, stage, meddra.disease.preferred.name, os, subject.dead.hospice, subject.age, line.1.group, line.1.therapy, prior.targeted_therapy, subject.metastatic.disease, irae.any, irae.severe)
  
# Merge data frames based on record_id and subject.id
dem_df <- left_join(df1, df3, by = c("record_id" = "subject.id"))

head(dem_df)
```

```{r}
# Create binary columns for each unique value in Class_1
unique_Revised_Classification <- unique(df1$Revised_Classification)

# Merge data frames based on record_id and subject.i
dem_df <- left_join(df1, df3, by = c("record_id" = "subject.id")) %>%

  mutate(

    Revised_Classification =

      case_when(Revised_Classification == "NA" ~ NA,

    TRUE ~ Revised_Classification))
head(dem_df)

dem_df <- dem_df %>%
  distinct() %>%
  mutate(Revised_Classification_present = 1) %>%
  pivot_wider(names_from = Revised_Classification, values_from = Revised_Classification_present, values_fill = 0)

colnames(dem_df)
```
```{r}
#restrict to pt cohort
dem_df <- dem_df %>%
  filter(!is.na(subject.dead.hospice)) %>%
  filter(!(stage == "UNK"))

nrow(dem_df)
```

Create corticosteroid column with corticosteroids unclear + corticosteroids systemic 

```{r}
dem_df <- dem_df %>%
  mutate(Corticosteroids = if_else(Corticosteroid_systemic == 1 | Corticosteroid_unclear == 1, 1, 0))
```


```{r}
#remove non systemic/supplement columns 
dem_df <- dem_df %>%
  select(- c("Vitamin", "Supplement", "Acid_suppression", "Drop_NA", "Cancer", "Anti_glaucoma_non_systemic", "Corticosteroid_non_systemic", "Anti_fungal_non_systemic", "Anesthetic_non_systemic", "Anti_histamine_non_systemic", "NA", "Antibiotics_non_systemic", "Cannabinoid", "Decongestant_non_systemic", "Mast_cell_stabilizer_non_systemic", "Cholinergic_non_systemic", "Bronchodilator"))
```

```{r}
colnames(dem_df)
```

```{r}
#restrict to drug classes w/ at least 50 patients
colsums <- colSums(dem_df[16:87])

ov50 <- names(colsums[colsums >= 50])

dem_df <-dem_df %>%
  select(1:16, all_of(ov50))

colnames(dem_df)
```
```{r}
# Save demographic dataframe
write.csv(
  dem_df,
  here::here("data", "processed", "RADIOHEAD_demographic_df.csv"),
  row.names = FALSE
)

message(sprintf("✓ Saved RADIOHEAD_demographic_df.csv with %d patients", nrow(dem_df)))
```

trim to just opioids 
```{r}
opioid_dems <- dem_df %>%
  select(1:13, "Opioid")
```


```{r session-info}
sessionInfo()
```



