---
title: "polypharmacy"
author: "Sammy Stone"
date: "2024-11-18"
output: html_document
---
```{r}
library(paletteer)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(openxlsx)
```

set wds: 

```{r}
wd <- list()
wd$main <- file.path("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD")
wd$data <- file.path(wd$main, "data")
wd$dataP <- file.path(wd$data, "processed")
wd$output <- file.path(wd$main, "output_sammy")
wd$script <- file.path(wd$main, "scripts_sammy")
```

Using sub_df and poly_df from radiohead_dfs

```{r}
#load sub_df
sub_df <- read_csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_multivariate_df.csv")
sub_df <- sub_df %>%
  select(-"...1")

#load polydf 
polydf <- read_csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_polypharmacy_df.csv")
polydf <- polydf %>%
  select(-"...1")

#load datmeta

dat_meta <- read.csv(file.path(wd$dataP, "round2/df.clean15.postMS_2024-08-22.csv"))

```
match patient counts 
```{r}
polydf <- polydf %>%
  filter(!stage=="UNK")
nrow(polydf)
```

mean and median calculations
```{r}
nrow(sub_df) 
print(mean(sub_df$polypharmacy))
print(median(sub_df$polypharmacy))
```


first- top drug classes: 
```{r}
top_drugs <- sub_df %>%
  select(12:41) #remove corticosteroids

colSums(top_drugs == 1)

#top ten are: Cardiovascular (601), HLD_TX (443), Anti-coagulants (424), Anti-emetic (389), Proton Pump Inhibitor(288), Anti-pyretics (344), Opioids(336), Anti-depressants (212), Antianxiety (199), Diabetes (196)

top_ten <- sub_df %>%
  select(record_id, Cardiovascular, HLD_Tx, Anti_coagulant, Anti_emetic, Anti_pyretic, Opioid, Proton_Pump_Inhibitor, Anti_depressant, Anti_anxiety, Diabetes) 

#check 
#View(top_ten)
#top_ten <- top_ten %>%
  ##select(-record_id)
```
```{r}
#using top ten with record id
drug_long <- top_ten %>%
  pivot_longer(cols = 2:11,  
               names_to = "drug", 
               values_to = "use") %>%
  filter(use == 1)  # Filter to only include patients who use the drug
View(drug_long)
```


```{r}
top_ten <- top_ten %>%
 select(-record_id)  

class_counts <- colSums(top_ten == 1)
#View(class_counts)

class_percent <- class_counts / 1024 * 100 
#iew(class_percent)
```


```{r}
# Combine counts and percentages into a dataframe
drug_summary <- data.frame(
  drug = names(class_counts),
  count = class_counts,
  percent = class_percent
)
#rename 
drug_summary <- drug_summary %>%
  mutate(drug = case_when(
    drug == "HLD_Tx" ~ "Hyperlipidemia Tx",
    drug == "Anti_coagulant" ~ "Anti-coagulants",
    drug == "Anti_emetic"  ~ "Anti-emetics",
    drug == "Anti_pyretic" ~ "Anti-pyretics",
    drug == "Anti_anxiety" ~ "Anti-anxiety",
    drug == "Proton_Pump_Inhibitor" ~ "Proton Pump Inhibitors",
    drug == "Anti_depressant" ~ "Anti-depressants",
    drug == "Opioid" ~ "Opioids",
    TRUE ~ drug
  ))

# Order by percent in increasing order and explicitly reorder 'drug' by 'percent'
drug_summary <- drug_summary %>%
  arrange(percent) %>%
  mutate(drug = factor(drug, levels = (unique(drug))))

# View the result to confirm
View(drug_summary)

# Check the levels of 'drug'
levels(drug_summary$drug)
```


#bar plot
```{r}
# Create the bar plot
bar <- ggplot(drug_summary, aes(x = drug, y = percent)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +  # Flip coordinates to make it horizontal
  theme_minimal() +
  labs(
    y = "Patient Exposure (%)", 
    x = NULL
  ) +
  theme(
    axis.text = element_text(size = 20, color = "black"),  # Axis tick labels
    axis.title = element_text(size = 25, color = "black") # Axis titles
  ) + scale_y_continuous(limits = c(0, 100))

print(bar)
```


```{r}
# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "drugclassbarplot.png")

# Open a PNG device
png(output_file_png, width = 8, height = 6, units = "in", res = 300)

# Print the plot
print(bar)

# Close the PNG device
dev.off()
```


###############################################################################

```{r}
median(polydf$polypharmacy)
```


# KM curve

```{r}
# Create survival object
surv_object <- Surv(time = polydf$os, event = polydf$subject.dead.hospice)   

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ polypharmacy_binary, data = polydf)

# Extract values
cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ polypharmacy_binary, data = polydf)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(fit, data = polydf, pval = TRUE, pval.size = 9,  risk.table.fontsize = 8, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("< 5 drug classes", "≥ 5 drug classes"),
                   palette = custom_colors,
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 25), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 25),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
    plot.title = element_text(size = 25),
  ) +
  xlab("Time (months)")
```


```{r}
# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_polypharmacy.png")

# Open a PNG device
png(output_file_png, width = 10, height = 9, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()
```

```{r}
print(plot)
```

KM curve but for lung subcohort
```{r}
#create lung subcohort 
lung.poly <- polydf %>%
  filter(meddra.disease.preferred.name == "Non-small cell lung cancer")
```

```{r}
# Create survival object
surv_object <- Surv(time = lung.poly$os, event = lung.poly$subject.dead.hospice)   

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ polypharmacy_binary, data = lung.poly)

# Extract values
cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ polypharmacy_binary, data = lung.poly)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(fit, data = lung.poly, pval = TRUE, pval.size = 9,  risk.table.fontsize = 8, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("< 5 drug classes", "≥ 5 drug classes"),
                   palette = custom_colors,
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 25), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 25),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
    plot.title = element_text(size = 25),
  ) +
  xlab("Time (months)")
```


```{r}
# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_lung_polypharmacy.png")

# Open a PNG device
png(output_file_png, width = 10, height = 9, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()
```

KM curve but for KM subcohort
```{r}
#create melanoma subcohort 
mel.poly <- polydf %>%
  filter(meddra.disease.preferred.name == "Malignant melanoma")
```

```{r}
# Create survival object
surv_object <- Surv(time = mel.poly$os, event = mel.poly$subject.dead.hospice)   

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ polypharmacy_binary, data = mel.poly)

# Extract values
cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ polypharmacy_binary, data = mel.poly)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(fit, data = mel.poly, pval = TRUE, pval.size = 9,  risk.table.fontsize = 8, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("< 5 drug classes", "≥ 5 drug classes"),
                   palette = custom_colors,
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 25), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 25),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
    plot.title = element_text(size = 25),
  ) +
  xlab("Time (months)")
```


```{r}
# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_melanoma_polypharmacy.png")

# Open a PNG device
png(output_file_png, width = 10, height = 9, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()
```

ICI only 
KM curve but for ICI subcohort
```{r}
#create melanoma subcohort 
ici.poly <- polydf %>%
  filter(line.1.therapy %in% c("pd1-pdl1", "ctla4+pd1-pdl1"))
```

```{r}
# Create survival object
surv_object <- Surv(time = ici.poly$os, event = ici.poly$subject.dead.hospice)   

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ polypharmacy_binary, data = ici.poly)

# Extract values
cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ polypharmacy_binary, data = ici.poly)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(fit, data = ici.poly, pval = TRUE, pval.size = 9,  risk.table.fontsize = 8, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("< 5 drug classes", "≥ 5 drug classes"),
                   palette = custom_colors,
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 25), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 25),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
    plot.title = element_text(size = 25),
  ) +
  xlab("Time (months)")
```


```{r}
# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_ici_polypharmacy.png")

# Open a PNG device
png(output_file_png, width = 10, height = 9, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()
```

KM but chemo only
KM curve but for chemo subcohort
```{r}
#create melanoma subcohort 
chemo.poly <- polydf %>%
  filter(line.1.therapy %in% c("chemo+pd1-pdl1", "chemo+ctla4+pd1-pdl1"))
```

```{r}
# Create survival object
surv_object <- Surv(time = chemo.poly$os, event = chemo.poly$subject.dead.hospice)   

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ polypharmacy_binary, data = chemo.poly)

# Extract values
cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ polypharmacy_binary, data = chemo.poly)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(fit, data = chemo.poly, pval = TRUE, pval.size = 9,  risk.table.fontsize = 8, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("< 5 drug classes", "≥ 5 drug classes"),
                   palette = custom_colors,
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 25), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 25),  # Axis tick labels
    axis.title = element_text(size = 25), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
    plot.title = element_text(size = 25),
  ) +
  xlab("Time (months)")
```


```{r}
# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_chemo_polypharmacy.png")

# Open a PNG device
png(output_file_png, width = 10, height = 9, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()
```





Chemo Only
4. Polypharmacy Stats 

```{r}
mean(polydf$polypharmacy)
median(polydf$polypharmacy)
```


Density Plots


```{r}
summary(polydf$subject.age)
```

Median/mean is 69. Let's create a new variable Age_Cat and separate patients into over/under 69 

```{r}
age_pdf <- polydf %>%
  mutate(agecat = ifelse(subject.age >= 69, "69+", "Under 69"))
```

sanity checks
```{r}
table(age_pdf$agecat)
```

```{r}
agecomp <- age_pdf %>%
  select(subject.age, agecat)

View(agecomp)
```



#Cancer Density Plots

```{r}
#plot kernel (density) plot
d1 <- polydf %>%
  ggplot( aes(x=polypharmacy)) +
    geom_density(color= "deepskyblue", size = 1.5) + #outline only, thicker line
    ggtitle("Pan-cancer") +
    theme_minimal() +
    theme(
      plot.title = element_text(size=10)
    )
d1
```

NSCLC 
```{r}
#filter by NSCLC 
nsclc_pdf <- polydf %>%
  filter(meddra.disease.preferred.name == "Non-small cell lung cancer")
```

Melanoma 
```{r}
#filter by melanoma
mel_pdf <- polydf %>%
  filter(meddra.disease.preferred.name == "Malignant melanoma")
```

```{r}
write.csv(mel_pdf, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/Top Ten Comparisons/melanomapolydf.csv", row.names = FALSE)
```

#Cancer Density Plots
Need to first combine dfs into a single dataset
Then can plot overlayed with ggplot2

```{r}
#Add grouping variable to each subdf 
polydf_pancancer <- polydf %>%
  mutate(group = "Pan-Cancer")
nsclc_pdf <- nsclc_pdf %>%
  mutate(group = "NSCLC")
mel_pdf <- mel_pdf %>%
  mutate(group = "Melanoma")
```


```{r}
#combine datasets
triple_df <- bind_rows(polydf_pancancer, nsclc_pdf, mel_pdf)
```

plot 

```{r}
# Plot the combined data
overlayed_density <- triple_df %>%
  ggplot(aes(x = polypharmacy, color = group)) +
  geom_density(size = 1.5, position = "identity") +
  scale_color_manual(values = c("Melanoma" = "orange",
                               "Pan-Cancer" = "darkgrey", 
                               "NSCLC" = "deepskyblue")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(), # remove legend title
    legend.text = element_text(size = 22),
     axis.text.x = element_text(size = 22), # increase x-axis text size
    axis.text.y = element_text(size = 22), # increase y-axis text size
    axis.title.x = element_text(size = 25), # increase x-axis title size
    axis.title.y = element_text(size = 25)  # increase y-axis title size
  ) +
  labs(
    x = "Number of Drug Classes", 
    y = "Density"
  ) +
  scale_y_continuous(limits = c(0, 0.2),
  expand = expansion(mult = c(0, 0.05))  # 5% headroom above tallest point
)

# Display the plot
plot(overlayed_density)
```

```{r}
ggsave(filename = file.path(wd$output, "/density plots/density plot by cancer.pdf"), plot = overlayed_density, device = "pdf", height = 6, width = 8) 
```


man-whitney wilcox

```{r}
cancer_anova <- aov(polypharmacy ~ group, 
                             data = triple_df)
summary(cancer_anova)
```
```{r}
TukeyHSD(cancer_anova)
```
melanoma medians differs significantly from others

now kruskal wallis (more appropriate)

```{r}
kruskal_results <- kruskal.test(polypharmacy ~ group, 
                             data = triple_df)

print(kruskal_results)
MKR_df <- data.frame(
  Statistic = kruskal_results$statistic,
  p_value = kruskal_results$p.value
)

```
```{r}
# Save the result as an Excel file
write.xlsx(MKR_df, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/density plots/pancancer_kruskal_test_result.xlsx")
```

#Chemo vs ICI only Density Plot

```{r}
ici_pdf <- polydf %>%
  filter(!line.1.therapy %in% c("chemo+pd1-pdl1", "chemo+ctla4+pd1-pdl1"))
```

```{r}
chemo_pdf <- polydf %>% 
  filter(line.1.therapy %in% c("chemo+pd1-pdl1", "chemo+ctla4+pd1-pdl1")) 

```

```{r}
ici_pdf <- ici_pdf %>%
  mutate(group = "ICI Only")
chemo_pdf <- chemo_pdf %>%
  mutate(group = "ICI + Chemo")
```

```{r}
write.csv(ici_pdf, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/Top Ten Comparisons/icipolydf.csv", row.names = FALSE)
write.csv(chemo_pdf, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/Top Ten Comparisons/chemopolydf.csv", row.names = FALSE)
```

```{r}
treatments_df <- bind_rows(ici_pdf, chemo_pdf)
treatments_df$group <- factor(treatments_df$group, levels = c("ICI Only", "ICI + Chemo"))
```

```{r}
# Plot the combined data
tx_density <- treatments_df %>%
  ggplot(aes(x = polypharmacy, color = group)) +
  geom_density(size = 1.5, position = "identity") +
  scale_color_manual(values = c("ICI Only" = "orange",
                               "ICI + Chemo" = "deepskyblue")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 22),
    legend.title = element_blank(), # remove legend title
    axis.text.x = element_text(size = 22), # increase x-axis text size
    axis.text.y = element_text(size = 22), # increase y-axis text size
    axis.title.x = element_text(size = 25), # increase x-axis title size
    axis.title.y = element_text(size = 25)  # increase y-axis title size
  ) +
  labs(
    x = "Number of Drug Classes", 
    y = "Density"
  ) +
scale_y_continuous(limits = c(0, 0.2), expand = expansion(mult = c(0, 0.05)))  # 5% headroom above tallest point

# Display the plot
plot(tx_density)
```

```{r}
ggsave(filename = file.path(wd$output, "/density plots/density plot by treatment.pdf"), plot = tx_density, device = "pdf", height = 6, width =8)
```

```{r}
#quick mean/median check
mean(chemo_pdf$polypharmacy)
median(chemo_pdf$polypharmacy)
mean(ici_pdf$polypharmacy)
median(ici_pdf$polypharmacy)

#data is right skewed, medians are better representations
```


stats test
```{r}
tx_wilcox <- wilcox.test(polypharmacy ~ group, 
                         data = treatments_df, 
                         exact = FALSE)
tx_wilcox

# Convert the result to a data frame
wilcox_results <- data.frame(
  Statistic = tx_wilcox$statistic,
  p_value = tx_wilcox$p.value
)

# Save the result as an Excel file
write.xlsx(wilcox_results, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/density plots/treatment_wilcox_test_result.xlsx")
```

#Density by Stage

create 3 dfs for each stage bin
```{r}
low_stage <- polydf %>%
  filter(stage %in% c("0", "i", "ii"))

stage3 <- polydf %>% 
  filter(stage == "iii")

stage4 <- polydf %>% 
  filter(stage == "iv")
```

add identifiers
```{r}
low_stage <- low_stage %>%
  mutate(group = "0-II")
stage3 <- stage3 %>%
  mutate(group = "III")
stage4 <- stage4 %>%
  mutate(group = "IV")
```

```{r}
stage_df <- bind_rows(low_stage, stage3, stage4)
```

plot 
```{r}
# Plot the combined data
overlayed_stage <- stage_df %>%
  ggplot(aes(x = polypharmacy, color = group)) +
  geom_density(size = 1.5, position = "identity") +
  scale_color_manual(values = c("0-II" = "darkgrey",
                               "III" = "deepskyblue", 
                               "IV" = "orange")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 22),
    legend.title = element_blank(), # remove legend title
    axis.text.x = element_text(size = 22), # increase x-axis text size
    axis.text.y = element_text(size = 22), # increase y-axis text size
    axis.title.x = element_text(size = 25), # increase x-axis title size
    axis.title.y = element_text(size = 25)  # increase y-axis title size
  ) +
  labs(
    x = "Number of Drug Classes", 
    y = "Density"
  ) +
scale_y_continuous(
  limits = c(0, 0.2),
  expand = expansion(mult = c(0, 0.05)))  # 5% headroom above tallest point


# Display the plot
plot(overlayed_stage)
```

```{r}
ggsave(filename = file.path(wd$output, "/density plots/density plot by stage.pdf"), plot = overlayed_stage, device = "pdf", height = 6, width =8)
```


stats 
```{r}
stage_anova <- aov(polypharmacy ~ group, 
                             data = stage_df)
summary(stage_anova)
```

```{r}
kruskal.test(polypharmacy ~ group, 
                             data = stage_df)


```

#By sex 

```{r}
#pull in subject.sex from dat_meta 
sex_only <- dat_meta %>%
  select(subject.id, subject.sex) %>%
  rename(record_id = subject.id)
```

```{r}
sex_df <- left_join(polydf, sex_only, by = "record_id")
```

new dfs
```{r}
female_df <- sex_df %>%
  filter(subject.sex == "female")

male_df <- sex_df %>%
  filter(subject.sex == "male")
```

label dfs
```{r}
female_df <- female_df %>%
  mutate(group = "Female")
male_df <- male_df %>%
  mutate(group = "Male")
```

```{r}
sex_plot_df <- bind_rows(female_df, male_df)
```

```{r}
# Plot the combined data
sex_density <- sex_plot_df %>%
  ggplot(aes(x = polypharmacy, color = group)) +
  geom_density(size = 1.5, position = "identity") +
  scale_color_manual(values = c("Female" = "orange",
                               "Male" = "deepskyblue")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 22),
    legend.title = element_blank(), #remove legend title
    axis.text.x = element_text(size = 22), # increase x-axis text size
    axis.text.y = element_text(size = 22), # increase y-axis text size
    axis.title.x = element_text(size = 25), # increase x-axis title size
    axis.title.y = element_text(size = 25)  # increase y-axis title size
    
  ) +
  labs(
    x = "Number of Drug Classes", 
    y = "Density"
  ) +
scale_y_continuous(
  limits = c(0, 0.2),
  expand = expansion(mult = c(0, 0.05)))  # 5% headroom above tallest point


# Display the plot
plot(sex_density)
```


```{r}
ggsave(filename = file.path(wd$output, "/density plots/density plot by sex.pdf"), plot = sex_density, device = "pdf", height = 6, width =8)
```




stats test
```{r}
sex_wilcox <- wilcox.test(polypharmacy ~ group, 
                         data = sex_plot_df, 
                         exact = FALSE)
sex_wilcox
```
#By Age 

Age binned as over/under 70

```{r}
age_catdf2 <- polydf %>%
  mutate(age_cat = case_when(
    subject.age < 70 ~ "Under 70",
    subject.age >= 70  ~ "70+"
  ))
```

sanity checks 
```{r}
unique(age_catdf2$age_cat)
table(age_catdf2$age_cat)
```

```{r}
un_70 <- age_catdf2 %>%
  filter(age_catdf2$age_cat == "Under 70")

plus_70 <- age_catdf2 %>%
  filter(age_catdf2$age_cat == "70+")

```

add identifiers
```{r}
un_70 <- un_70 %>%
  mutate(group = "Under 70")
plus_70 <- plus_70 %>%
  mutate(group = "70+")

```

```{r}
write.csv(un_70, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/Top Ten Comparisons/under70polydf.csv", row.names = FALSE)
write.csv(plus_70, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/Top Ten Comparisons/over70polydf.csv", row.names = FALSE)
```

```{r}
age_grouped2 <- bind_rows(un_70, plus_70)
age_grouped2$group <- factor(age_grouped2$group, levels = c("Under 70", "70+"))

```

```{r}
# Plot the combined data
age_density2 <- age_grouped2 %>%
  ggplot(aes(x = polypharmacy, color = group)) +
  geom_density(size = 1.5, position = "identity") +
  scale_color_manual(
    values = c("Under 70" = "orange", "70+" = "deepskyblue"),
    labels = c("Under 70", "70+"))  +  # Set legend label order
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 22),
    legend.title = element_blank(), # remove legend title
    axis.text.x = element_text(size = 22), # increase x-axis text size
    axis.text.y = element_text(size = 22), # increase y-axis text size
    axis.title.x = element_text(size = 25), # increase x-axis title size
    axis.title.y = element_text(size = 25)  # increase y-axis title size
  ) +
  labs(
    x = "Number of Drug Classes", 
    y = "Density"
  ) +
scale_y_continuous(
  limits = c(0, 0.2),
  expand = expansion(mult = c(0, 0.05))  # 5% headroom above tallest point
)

# Display the plot
plot(age_density2)
```
```{r}
#ggsave(filename = file.path(wd$output, "/density plots/density plot by age2.pdf"), plot = age_density2, device = "pdf", height = 6, width =8)
```

```{r}
age_wilcox2 <- wilcox.test(polypharmacy ~ group, 
                         data = age_grouped2, 
                         exact = FALSE)
age_wilcox2 


age_wilcox2_results <- data.frame(
  Statistic = age_wilcox2$statistic,
  p_value = age_wilcox2$p.value
)

#write.xlsx(age_wilcox2_results, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output_sammy/density plots/age_wilcox_test_result.xlsx")

```


#Polypharmacy Frequency Plot

```{r}
poly_pie_2 <- sub_df %>%
  select(record_id, polypharmacy)
```

```{r}
poly_sum2 <- poly_pie_2 %>%
  select(record_id, polypharmacy) %>%
  group_by(polypharmacy) %>%
  summarize(count = n(), .groups = "drop")

```

```{r}
poly_freq2 <- poly_sum2 %>%
  mutate(freq = (count / 1024) *100)
```

```{r}
# Ensure polypharmacy is a factor
poly_freq2 <- poly_freq2 %>%
  mutate(polypharmacy = as.factor(polypharmacy))
```



```{r}
custom_colors <- c("darkorange3", "darkorange", "orange2", "orange1", "orange", "skyblue", "lightskyblue", "skyblue1", "skyblue2", "deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3", "dodgerblue", "dodgerblue2", "dodgerblue3", "royalblue3", "dodgerblue4", "royalblue4", "midnightblue")
```


```{r}
freq <- ggplot(poly_freq2, aes(x = polypharmacy, y = freq)) +
  geom_segment(aes(x = polypharmacy, xend = polypharmacy, y = 0, yend = freq), color = "darkgrey", size = 0.5) +  # Create thicker lines
  geom_point(aes(color = polypharmacy), size = 5) +  # Create the points
 scale_color_manual(values = custom_colors) +
  labs(x = "Polypharmacy (Continuous)", y = "Frequency (%)", color = "Polypharmacy") +  # Adjust axis and legend titles
  theme_minimal() + 
  theme(legend.position = "none") + 
theme(axis.text = element_text(size = 18, color = "black"), 
        axis.title.x = element_text(size = 18, color = "black", margin = margin(t = 20)),  # Move x-axis title farther from plot
        axis.title.y = element_text(size = 18, color = "black", margin = margin(r = 20)),  # Move y-axis title farther from plot
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank(),  # Remove minor grid lines
        axis.line = element_line(color = "darkgrey")) +  # Keep axis lines
  ylim(0, 15)  # Set the y-axis limit
freq

ggsave(filename = file.path(wd$output, "polypharmacylollipop_chart.png"), plot = freq, width = 8, height = 6, dpi = 300)
```

