---
title: "02 - Polypharmacy Analysis"
author: "Sammy Stone"
date: "2026-1-26"
output: html_document
---

# Polypharmacy Analysis (Figure 1)

# This script analyzes polypharmacy patterns in the cohort:
# 1. Kaplan-Meier survival curves by polypharmacy status (≥5 vs <5 drug classes)
# 2. Polypharmacy frequency distribution (lollipop chart)

# Key Outputs:
# - Figure 1B KM_polypharmacy.png (Kaplan-Meier curve)
# - Figure 1A polypharmacy_lollipop_chart.png (frequency distribution)

# ==============================================================================
# Setup
# ==============================================================================

```{r setup, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(survival)
library(survminer)

# Load project configuration
source(here::here("R", "config.R"))
```

# ==============================================================================
# Load Data
# ==============================================================================

```{r load-data}
# Load multivariate dataframe
sub_df <- read.csv(paths$multivariate_df)

# Load polypharmacy dataframe
polydf <- read.csv(paths$polypharmacy_df)

# Filter to match cohort (exclude UNK stage)
polydf <- polydf %>%
  filter(stage != "UNK")

message(sprintf("✓ Loaded %d patients for analysis", nrow(polydf)))
message(sprintf("  Mean polypharmacy: %.2f", mean(polydf$polypharmacy)))
message(sprintf("  Median polypharmacy: %.0f", median(polydf$polypharmacy)))
```

# ==============================================================================
# Kaplan-Meier Survival Analysis by Polypharmacy Status
# ==============================================================================
**Fig 1B**
```{r km-polypharmacy, fig.width=10, fig.height=9}
# Create survival object
surv_object <- Surv(time = polydf$os, event = polydf$subject.dead.hospice)

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ polypharmacy_binary, data = polydf)

# Cox model for p-value
cox_model <- coxph(Surv(os, subject.dead.hospice) ~ polypharmacy_binary, data = polydf)

# Define custom colors
custom_colors <- c("orange", "deepskyblue")

# Create the plot
km_plot <- ggsurvplot(
  fit,
  data = polydf,
  pval = TRUE,
  pval.size = 9,
  risk.table = TRUE,
  risk.table.fontsize = 8,
  conf.int = TRUE,
  legend.labs = c("< 5 drug classes", "≥ 5 drug classes"),
  palette = custom_colors,
  legend.title = ""
)

# Customize main plot
km_plot$plot <- km_plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 25),
    legend.text = element_text(size = 25),
    legend.title = element_text(size = 25)
  ) +
  xlab("Time (months)")

# Customize risk table
km_plot$table <- km_plot$table +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 25),
    axis.title = element_text(size = 25),
    legend.text = element_text(size = 25),
    plot.title = element_text(size = 25)
  ) +
  xlab("Time (months)")

# Display plot
print(km_plot)

# Save plot
png(
  here::here("output", "figures", "Fig_1B.png"),
  width = 10,
  height = 9,
  units = "in",
  res = 300
)
print(km_plot) 
dev.off()

message("✓ Saved Fig_1B.png")
```

# ==============================================================================
# Polypharmacy Frequency Distribution (Lollipop Chart)
# ==============================================================================
**Fig 1A**
```{r polypharmacy-frequency}
# Calculate frequency distribution
poly_freq <- sub_df %>%
  select(record_id, polypharmacy) %>%
  group_by(polypharmacy) %>%
  summarize(count = n(), .groups = "drop") %>%
  mutate(
    freq = (count / 1024) * 100,
    polypharmacy = as.factor(polypharmacy)
  )

# Define custom color gradient
custom_colors <- c(
  "darkorange3", "darkorange", "orange2", "orange1", "orange",
  "skyblue", "lightskyblue", "skyblue1", "skyblue2",
  "deepskyblue", "deepskyblue1", "deepskyblue2", "deepskyblue3",
  "dodgerblue", "dodgerblue2", "dodgerblue3", "royalblue3",
  "dodgerblue4", "royalblue4", "midnightblue"
)

# Create lollipop chart
lollipop_plot <- ggplot(poly_freq, aes(x = polypharmacy, y = freq)) +
  geom_segment(
    aes(x = polypharmacy, xend = polypharmacy, y = 0, yend = freq),
    color = "darkgrey",
    linewidth = 0.5
  ) +
  geom_point(aes(color = polypharmacy), size = 5) +
  scale_color_manual(values = custom_colors) +
  labs(
    x = "Number of Drug Classes",
    y = "Frequency (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 18, color = "black"),
    axis.title.x = element_text(size = 18, color = "black", margin = margin(t = 20)),
    axis.title.y = element_text(size = 18, color = "black", margin = margin(r = 20)),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "darkgrey")
  ) +
  ylim(0, 15)

# Display plot
print(lollipop_plot)

# Save plot
png(
  here::here("output", "figures", "Fig_1A.png"),
  width = 8,
  height = 6,
  units = "in",
  res = 300
)
print(lollipop_plot) 
dev.off()
message("✓ Saved Fig_1A.png") 
```



# ==============================================================================
# Summary Statistics
# ==============================================================================

```{r summary-stats}
# Polypharmacy distribution
poly_summary <- sub_df %>%
  summarize(
    n = n(),
    mean = mean(polypharmacy),
    median = median(polypharmacy),
    sd = sd(polypharmacy),
    min = min(polypharmacy),
    max = max(polypharmacy)
  )

print(poly_summary)

# Polypharmacy binary breakdown
poly_binary <- polydf %>%
  group_by(polypharmacy_binary) %>%
  summarize(
    n = n(),
    percent = n() / nrow(polydf) * 100
  )

print(poly_binary)

message("\n=== Polypharmacy Analysis Complete ===")
message(sprintf("Total patients: %d", nrow(polydf)))
message(sprintf("Patients with ≥5 drug classes: %d (%.1f%%)",
                sum(polydf$polypharmacy_binary),
                sum(polydf$polypharmacy_binary) / nrow(polydf) * 100))
```

```{r session-info}
sessionInfo()
```
