---
title: "03 - Multivariate Analysis"
author: "Sammy Stone"
date: "2024-07-31"
output:
  word_document: default
  html_document: default
---

# Multivariate and Univariate Drug Class Analysis
#
# This script performs Cox proportional hazards regression to assess the association
# between drug class exposure and overall survival, controlling for clinical covariates.
#
# Key Analyses:
# 1. Univariate and multivariate Cox regression for all drug classes (≥50 patients)
# 2. Combined multivariate model with significant drug classes
# 3. Opioid + palliative bundle interaction analysis
# 4. Forest plots for visualization
#
# Key Outputs:
# - Merged univariate/multivariate table (Excel)
# - Forest_plot_1.png (significant drug classes)
# - Forest_plot_2.png (opioid-palliative interaction)
# - palliativebundledf.csv (for use in subsequent analyses)

# ==============================================================================
# Setup
# ==============================================================================

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(survival)
library(gtsummary)
library(survminer)
library(broom)
library(openxlsx)
library(patchwork)
library(car)

# Load project configuration
source(here::here("R", "config.R"))
```

# ==============================================================================
# Load Data
# ==============================================================================

```{r load-data}
sub_df <- read.csv(paths$multivariate_df)

# Remove unnecessary columns
dat <- sub_df %>%
  select(-any_of(c("X", "Bronchodilator")))

# Verify no NA patients
d1 <- dat %>% 
  filter(!is.na(subject.dead.hospice))

message(sprintf("✓ Loaded %d patients for multivariate analysis", nrow(d1)))
message(sprintf("✓ Analyzing %d drug classes", length(names(d1)[11:41])))
```

# ==============================================================================
# Step 1: Multivariate Cox Regression (All Drug Classes)
# ==============================================================================
**Table 1**
```{r multivariate-analysis}
# Relevel covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")

# Define baseline covariates
base_covariates <- c("subject.age", "prior.targeted_therapy", "subject.metastatic.disease", 
                     "Treatment", "cancer_agreggate", "stage")

# Get drug class columns
drug_class_columns <- names(d1)[11:41]

# Initialize results list
cox_results <- list()

# Loop through each drug class
for (drug_class in drug_class_columns) {
  
  # Fit Cox model
  formula_str <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                        paste(base_covariates, collapse = " + "), " + ", drug_class)
  
  fit <- coxph(as.formula(formula_str), data = d1)
  
  # Create summary table (show only drug class, suppress covariates)
  tbl <- tbl_regression(fit, exponentiate = TRUE) %>%
    modify_table_body(~ .x %>% filter(variable == drug_class)) %>%
    modify_header(label = paste0("Drug Class: ", drug_class))
  
  # Add patient count
  count <- sum(d1[[drug_class]] == 1, na.rm = TRUE)
  
  tbl <- tbl %>%
    modify_table_body(~ .x %>% mutate(Count = count)) %>%
    modify_header(Count = "**Count**")
  
  cox_results[[drug_class]] <- tbl
}

# Sort by patient count
drug_counts <- data.frame(
  Drug = drug_class_columns, 
  Count = sapply(drug_class_columns, function(dc) sum(d1[[dc]] == 1, na.rm = TRUE))
)
drug_counts <- drug_counts[order(-drug_counts$Count), ]

cox_results <- cox_results[match(drug_counts$Drug, names(cox_results))]

# Stack results
final_table <- tbl_stack(cox_results) %>%
  modify_header(label = "Drug Class") %>%
  bold_p(t = 0.05)

print(final_table)
```

## Add Benjamini-Hochberg Correction

```{r bh-correction}
# Extract p-values
p_values <- final_table$table_body$p.value

# Perform BH correction
q_values <- p.adjust(p_values, method = "BH")
q_values <- style_pvalue(q_values)

# Add q-values to table
final_table <- final_table %>%
  modify_table_body(~ .x %>% mutate(`Q-value` = q_values)) %>%
  modify_header(`Q-value` = "**q-value**")

print(final_table)
```

# ==============================================================================
# Step 2: Univariate Cox Regression (All Drug Classes)
# ==============================================================================

```{r univariate-analysis}
# Initialize results list
uv_results <- list()

# Loop through each drug class
for (drug_class in drug_class_columns) {
  
  # Fit univariate Cox model
  formula_uv <- as.formula(paste0("Surv(os, subject.dead.hospice) ~ `", drug_class, "`"))
  fit_uv <- coxph(formula_uv, data = d1)
  
  # Create summary table
  tbl_uv <- tbl_regression(fit_uv, exponentiate = TRUE) %>%
    modify_header(label = "Covariate") %>%
    modify_table_body(~ .x %>% mutate(variable = drug_class))
  
  uv_results[[drug_class]] <- tbl_uv
}

# Reorder by patient count
uv_results_ordered <- uv_results[match(drug_counts$Drug, names(uv_results))]

# Stack results
univariate_table <- tbl_stack(uv_results_ordered) %>%
  bold_p(t = 0.05)

print(univariate_table)
```

# ==============================================================================
# Step 3: Merge Univariate and Multivariate Tables
# ==============================================================================

```{r merge-tables}
# Merge tables with headers
merged_table <- tbl_merge(
  tbls = list(univariate_table, final_table),
  tab_spanner = c("**Univariate**", "**Multivariate**")
)

# Display as formatted table
merged_table %>% 
  as_gt() %>% 
  gt::tab_options(table.font.names = "Arial")

# Save merged table to Excel
merged_table %>%
  as_hux_xlsx(here::here("output", "tables", "Table_1.xlsx"))

message("✓ Saved Table_1.xlsx")
```

# ==============================================================================
# Step 4: Combined Multivariate Model (Significant Drugs)
# ==============================================================================
**Table S2**
```{r combined-multivariate}
# Select significant drug classes and covariates
d1_combined <- dat %>%
  select(record_id, os, subject.dead.hospice, subject.age, prior.targeted_therapy, 
         subject.metastatic.disease, Treatment, cancer_agreggate, stage, 
         Opioid, Corticosteroid_systemic, Beta_Blockers, Diuretic, 
         Laxative, Anti_emetic, Benzodiazepines)

# Relevel covariates
d1_combined$Treatment <- relevel(factor(d1_combined$Treatment), ref = "pd1-pdl1")
d1_combined$cancer_agreggate <- relevel(factor(d1_combined$cancer_agreggate), 
                                        ref = "Non-small cell lung cancer")
d1_combined$stage <- relevel(factor(d1_combined$stage), ref = "iv")

# Fit combined Cox model
formula_combined <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                          paste(colnames(d1_combined)[-(1:3)], collapse = " + "))

fit_combined <- coxph(as.formula(formula_combined), data = d1_combined)

# Create summary tables
gt_table <- tbl_regression(fit_combined)
gt_table2 <- tbl_regression(fit_combined, exponentiate = TRUE)

print(gt_table)
print(gt_table2)

# Save combined multivariate tables to Excel
gt_table %>%
  as_hux_xlsx(here::here("output", "tables", "Table_S2_Log.xlsx"))

gt_table2 %>%
  as_hux_xlsx(here::here("output", "tables", "Table_S2.xlsx"))

message("✓ Saved Table S2 tables")

# Check VIF for multicollinearity
vif_model <- lm(Opioid ~ polypharmacy + Corticosteroid_systemic + Beta_Blockers + 
                Diuretic + Laxative + Anti_emetic + Benzodiazepines + 
                prior.targeted_therapy + stage + Treatment, data = sub_df)

message("\n=== VIF Analysis (Multicollinearity Check) ===")
print(vif(vif_model))

# Check correlation
message(sprintf("\nCorrelation (Opioid-Polypharmacy): %.3f", 
                cor(sub_df$Opioid, sub_df$polypharmacy)))
```

# ==============================================================================
# Step 5: Forest Plot - Significant Drug Classes
# ==============================================================================
**Fig 2A**
```{r forest-plot-1, fig.width=12, fig.height=6}
# Convert model output to tibble
forest_data <- tidy(fit_combined, conf.int = TRUE)

# Filter to drug classes only
c_forest <- forest_data %>%
  filter(term %in% c("Opioid", "Corticosteroid_systemic", "Beta_Blockers", 
                     "Diuretic", "Laxative", "Anti_emetic", "Benzodiazepines"))

# Rename for visualization
c_forest <- c_forest %>%
  mutate(term = case_when(
    term == "Opioid" ~ "Opioids",
    term == "Anti_emetic" ~ "Anti-emetics",
    term == "Diuretic" ~ "Diuretics",
    term == "Corticosteroid_systemic" ~ "Corticosteroids (systemic)",
    term == "Beta_Blockers" ~ "Beta-blockers",
    term == "Laxative" ~ "Laxatives",
    term == "Benzodiazepines" ~ "Benzodiazepines",
    TRUE ~ term
  )) %>%
  arrange(desc(estimate)) %>%
  mutate(term = factor(term, levels = rev(term)))

# Create forest plot
forest_plot <- ggplot(c_forest, aes(x = estimate, y = term)) +
  geom_point(size = 3, color = "deepskyblue") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, 
                 color = "deepskyblue") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
  scale_x_continuous(
    name = "Log Hazard Ratio",
    breaks = scales::pretty_breaks(),
    limits = c(-0.2, 1)
  ) +
  scale_y_discrete(name = "") +
  theme_minimal(base_size = 22) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 25, color = "black"),
    axis.text.x = element_text(size = 25)
  )

# Create p-value plot
p_plot <- ggplot(c_forest, aes(x = 0, y = term,
                               label = ifelse(p.value < 0.001, "p < 0.001", 
                                            sprintf("p = %.3f", p.value)))) +
  geom_text(aes(fontface = ifelse(p.value < 0.05, "bold", "plain")),
            hjust = 0, size = 8) +
  theme_void() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(0, 0, 0, 10)
  )

# Combine plots
layout <- c(
  area(t = 0, l = 0, b = 30, r = 14),
  area(t = 0, l = 12, b = 30, r = 18)
)

combined_forest <- forest_plot + p_plot + plot_layout(design = layout)
print(combined_forest)

# Save
ggsave(
  here::here("output", "figures", "Fig_2A.png"),
  plot = combined_forest,
  width = 12,
  height = 6,
  dpi = 300
)

message("✓ Saved Fig_2A.png")
```

# ==============================================================================
# Step 6: Opioid + Palliative Bundle Interaction Analysis
# ==============================================================================
**Table S4**
## Create Palliative Bundle Variable

```{r create-palliative-bundle}
# Define palliative drugs
paldrugs <- c("Laxative", "Benzodiazepines", "Anti_emetic")

# Create palliative bundle indicator
pall_df <- sub_df %>%
  mutate(palliative_drugs = ifelse(rowSums(across(all_of(paldrugs))) > 0, 1, 0))

# Save for use in other scripts
write.csv(
  pall_df,
  paths$palliative_bundle_df,
  row.names = FALSE
)

message("✓ Saved palliativebundledf.csv")
message(sprintf("  Patients with palliative bundle: %d", sum(pall_df$palliative_drugs)))
message(sprintf("  Patients with opioids: %d", sum(pall_df$Opioid)))
message(sprintf("  Patients with both: %d", 
                sum(pall_df$Opioid == 1 & pall_df$palliative_drugs == 1)))
```

## Multivariate Model with Interaction Term

```{r opioid-interaction}
# Prepare data
d1_interaction <- pall_df %>%
  select(record_id, os, subject.dead.hospice, subject.age, prior.targeted_therapy,
         subject.metastatic.disease, Treatment, cancer_agreggate, stage,
         Opioid, palliative_drugs)

# Relevel covariates
d1_interaction$Treatment <- relevel(factor(d1_interaction$Treatment), ref = "pd1-pdl1")
d1_interaction$cancer_agreggate <- relevel(factor(d1_interaction$cancer_agreggate),
                                           ref = "Non-small cell lung cancer")
d1_interaction$stage <- relevel(factor(d1_interaction$stage), ref = "iv")

# Fit model with interaction
formula_interaction <- paste0("Surv(os, subject.dead.hospice) ~ ",
                             paste(colnames(d1_interaction)[-(1:3)], collapse = " + "),
                             " + Opioid:palliative_drugs")

fit_interaction <- coxph(as.formula(formula_interaction), data = d1_interaction)

# Summary tables
gt_table_int <- tbl_regression(fit_interaction)
gt_table_int2 <- tbl_regression(fit_interaction, exponentiate = TRUE)

print(gt_table_int)
print(gt_table_int2)

# Save interaction model tables to Excel
gt_table_int %>%
  as_hux_xlsx(here::here("output", "tables", "Table_S4_Log.xlsx"))

gt_table_int2 %>%
  as_hux_xlsx(here::here("output", "tables", "Table_S4.xlsx"))

message("✓ Saved Table S4 tables")

# Calculate counts for interaction terms
interaction_counts <- c(
  Opioid = sum(d1_interaction$Opioid == 1, na.rm = TRUE),
  Palliative_drugs = sum(d1_interaction$palliative_drugs == 1, na.rm = TRUE),
  Interaction = sum(d1_interaction$Opioid == 1 & d1_interaction$palliative_drugs == 1, 
                   na.rm = TRUE)
)

message("\n=== Interaction Term Counts ===")
print(interaction_counts)
```

# ==============================================================================
# Step 7: Forest Plot - Opioid-Palliative Interaction
# ==============================================================================
**Figure 3B**
```{r forest-plot-2, fig.width=12, fig.height=6}
# Convert model output to tibble
forest_data_int <- tidy(fit_interaction, conf.int = TRUE)

# Filter to variables of interest
o_forest <- forest_data_int %>%
  filter(term %in% c("Opioid", "palliative_drugs", "Opioid:palliative_drugs"))

# Rename for visualization
o_forest <- o_forest %>%
  mutate(term = case_when(
    term == "Opioid" ~ "Opioids",
    term == "palliative_drugs" ~ "Palliative Bundle",
    term == "Opioid:palliative_drugs" ~ "Opioids * Palliative Bundle",
    TRUE ~ term
  )) %>%
  arrange(desc(estimate)) %>%
  mutate(term = factor(term, levels = rev(term)))

# Create forest plot
forest_plot2 <- ggplot(o_forest, aes(x = estimate, y = term)) +
  geom_point(size = 3, color = "orange") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, 
                 color = "orange") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
  scale_x_continuous(
    name = "Log Hazard Ratio",
    breaks = scales::pretty_breaks(),
    limits = c(-0.5, 1)
  ) +
  scale_y_discrete(name = "") +
  theme_minimal(base_size = 22) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 25, color = "black"),
    axis.text.x = element_text(size = 25)
  )

# Create p-value plot
p_plot2 <- ggplot(o_forest, aes(x = 0, y = term,
                                label = ifelse(p.value < 0.001, "p < 0.001",
                                             sprintf("p = %.3f", p.value)))) +
  geom_text(aes(fontface = ifelse(p.value < 0.05, "bold", "plain")),
            hjust = 0, size = 8) +
  theme_void() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(0, 0, 0, 10)
  )

# Combine plots
layout2 <- c(
  area(t = 0, l = 0, b = 20, r = 12),
  area(t = 0, l = 11, b = 20, r = 16)
)

combined_forest2 <- forest_plot2 + p_plot2 + plot_layout(design = layout2)
print(combined_forest2)

# Save
ggsave(
  here::here("output", "figures", "Fig_3B.png"),
  plot = combined_forest2,
  width = 12,
  height = 6,
  dpi = 300
)

message("✓ Saved Fig_3B .png")
```

# ==============================================================================
# Step 8: Kaplan-Meier Survival Curves
# ==============================================================================
**Figure 2C**
## Opioid Exposure KM Curve
```{r km-opioid, fig.width=10, fig.height=8}
# Create survival object
surv_object <- Surv(time = sub_df$os, event = sub_df$subject.dead.hospice)

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ Opioid, data = sub_df)

# Cox model for p-value
cox_model <- coxph(Surv(os, subject.dead.hospice) ~ Opioid, data = sub_df)

# Define custom colors
custom_colors <- c("orange", "deepskyblue")

# Create the plot
km_opioid <- ggsurvplot(
  fit,
  data = sub_df,
  pval = TRUE,
  pval.size = 9,
  risk.table = TRUE,
  risk.table.fontsize = 8,
  conf.int = TRUE,
  legend.labs = c("No Opioid Exposure", "Opioid Exposure"),
  palette = custom_colors,
  legend.title = ""
)

# Customize main plot
km_opioid$plot <- km_opioid$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.6, 0.2),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 25),
    legend.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Customize risk table
km_opioid$table <- km_opioid$table +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 20),
    plot.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Display plot
print(km_opioid)

# Save plot using png() device to preserve risk table
png(
  here::here("output", "figures", "Fig_2C.png"),
  width = 10,
  height = 8,
  units = "in",
  res = 300
)
print(km_opioid)
dev.off()

message("✓ Saved Fig_2C.png")
```

**Fig 2D**
## Beta-Blocker Exposure KM Curve
```{r km-beta-blocker, fig.width=10, fig.height=8}
# Create survival object
surv_object <- Surv(time = sub_df$os, event = sub_df$subject.dead.hospice)

# Fit Kaplan-Meier curve
fit <- survfit(surv_object ~ Beta_Blockers, data = sub_df)

# Cox model for p-value
cox_model <- coxph(Surv(os, subject.dead.hospice) ~ Beta_Blockers, data = sub_df)

# Define custom colors
custom_colors <- c("orange", "deepskyblue")

# Create the plot
km_beta <- ggsurvplot(
  fit,
  data = sub_df,
  pval = TRUE,
  pval.size = 9,
  risk.table = TRUE,
  risk.table.fontsize = 8,
  conf.int = TRUE,
  legend.labs = c("No Beta-Blocker Exposure", "Beta-Blocker Exposure"),
  palette = custom_colors,
  legend.title = ""
)

# Customize main plot
km_beta$plot <- km_beta$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.6, 0.2),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 25),
    legend.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Customize risk table
km_beta$table <- km_beta$table +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 20),
    plot.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Display plot
print(km_beta)

# Save plot using png() device to preserve risk table
png(
  here::here("output", "figures", "Fig_2D.png"),
  width = 10,
  height = 8,
  units = "in",
  res = 300
)
print(km_beta)
dev.off()

message("✓ Saved Fig_2D.png")
```
# ==============================================================================
# Summary
# ==============================================================================
```{r summary}
message("\n=== Multivariate Analysis Complete ===")
message(sprintf("Analyzed %d drug classes", length(drug_class_columns)))
message(sprintf("Significant in multivariate (p<0.05): %d", 
                sum(final_table$table_body$p.value < 0.05, na.rm = TRUE)))
message(sprintf("Significant after BH correction (q<0.05): %d",
                sum(as.numeric(gsub("[^0-9.]", "", q_values)) < 0.05, na.rm = TRUE)))

message("\n=== Output Files ===")
message("\nTables:")
message("  - output/tables/Table_1.xlsx")
message("  - output/tables/Table_S2.xlsx")
message("  - output/tables/Table_S2_Log.xlsx")
message("  - output/tables/Table_S4.xlsx")
message("  - output/tables/Table_S4_Log.xlsx")

message("\nFigures:")
message("  - output/figures/Figure_2A.png")
message("  - output/figures/Fig_2C.png")
message("  - output/figures/Fig_2D.png")
message("  - output/figures/Figure_3B.png")


```
```{r session-info}
sessionInfo()
```

```{r session-info}
sessionInfo()
```
