---
title: "multivariate and univariate RADIOHEAD drug analysis"
author: "Sammy Stone"
date: "2024-07-31"
output:
  word_document: default
  html_document: default
---

Multivariate and Univariate analysis of top drug classes on radiohead pt survival

Goal: Create multivariate tables for manuscript by using gtsummary to tabulate multivariate and univariate regresson analyses 

```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(gtsummary)
library(readr)
library(survminer)
library(ggplot2)
library(readxl)
library(broom)
library(openxlsx)
library(writexl)
library(patchwork)


```

#### Directories

```{r}
wd <- list()
wd$main <- "/Users/samanthastone/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD"
wd$data <- file.path(wd$main, "data")
wd$dataP <- file.path(wd$data, "processed")
wd$output <- file.path(wd$main, "output_sammy")
wd$script <- file.path(wd$main, "scripts_sammy")
```

#Load
```{r}
sub_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_multivariate_df.csv")
```

#(50+) w/ sub classes of interest and stage and polypharmacy

using sub_df from Radiohead_dfs.Rmd (file with polypharmacy, and drugs + sub classes of interest with at least 50 patients)

```{r}
dat <- sub_df %>%
  select(-"X", -"Bronchodilator")

#verify no NA patients 
d1 = dat %>% filter(!is.na(subject.dead.hospice))
```


##Step 1: Multivariate Table

```{r}
# Relevel the covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")

# List of baseline covariates
base_covariates <- c("subject.age", "prior.targeted_therapy", "subject.metastatic.disease", "Treatment", "cancer_agreggate", "stage")

# Initialize a list to store the results for each drug class
cox_results <- list()

# Loop through drug classes 
drug_class_columns <- names(d1)[11:41]

for (drug_class in drug_class_columns) {
  
   # Fit the Cox proportional hazards model with covariates and the drug class
  formula_str <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                        paste(base_covariates, collapse = " + "), " + ", drug_class)
  
  fit <- coxph(as.formula(formula_str), data = d1)
  
  # Summarize only the drug class in the table output (suppress the base covariates)
  tbl <- tbl_regression(fit, exponentiate = TRUE) %>%
    modify_table_body(~ .x %>%
      filter(variable == drug_class)) %>%
    modify_header(label = paste0("Drug Class: ", drug_class))
  
# Count the number of patients with a value of 1 in the covariate column
   count <- sum(d1[[drug_class]] == 1, na.rm = TRUE)
   
# Add the count as a new column in the table
  tbl <- tbl %>%
    modify_table_body(~ .x %>%
      mutate(Count = count)) %>%
    modify_header(Count = "**Count**")  # Add the count to the header
   
  # Store the results
  cox_results[[drug_class]] <- tbl
}

##Sorting by counts##
# Create a dataframe for sorting based on counts
drug_counts <- data.frame(Drug = drug_class_columns, Count = sapply(drug_class_columns, function(dc) sum(d1[[dc]] == 1, na.rm = TRUE)))

# Sort the drug counts in descending order
drug_counts <- drug_counts[order(-drug_counts$Count),]

# Reorder the cox_results based on sorted drug counts
cox_results <- cox_results[match(drug_counts$Drug, names(cox_results))] 

# Create a stacked gtsummary table
final_table <- tbl_stack(cox_results) %>%
   modify_header(label = "Drug Class")

#bold the p's
final_table <- final_table %>%
  bold_p(t = 0.05)  

# Print the final table
final_table

##the count here is a bit redundant should streamline
```

###Add BH correction
```{r}
#Extract p-values 
p_values <- final_table$table_body$p.value 

#Perform BH correction
q_values <- p.adjust(p_values, method = "BH")

# Format q-values using gtsummary's style_pvalue()
q_values <- style_pvalue(q_values)

#Add to table 
final_table <- final_table %>% 
  modify_table_body(~ .x %>%
  mutate(`Q-value` = q_values)) %>%
  modify_header(`Q-value` = "**q-value**")

print(final_table)
  
```
```{r}
ov5 <- sub_df %>%
  filter(polypharmacy >= 5)
nrow(ov5)
```

##Step 2: Univariate Table
```{r}
# Initialize a list to store the results for each drug class in univariate analysis
uv_results <- list()

# Loop through each drug class for univariate analysis
for (drug_class in drug_class_columns) {
  
  # Fit a univariate Cox proportional hazards model for each drug class
  formula_uv <- as.formula(paste0("Surv(os, subject.dead.hospice) ~ `", drug_class, "`"))
  fit_uv <- coxph(formula_uv, data = d1)
  
  # Create a table summarizing the results for the univariate analysis
  tbl_uv <- tbl_regression(fit_uv, exponentiate = TRUE) %>%
     modify_header(label = "Covariate") %>%
    modify_table_body(~ .x %>% mutate(variable = drug_class))  # Ensure the correct drug class name appears
  
  # Store the univariate result
  uv_results[[drug_class]] <- tbl_uv
}
# Use the drug counts to reorder the univariate results
uv_results_ordered <- uv_results[match(drug_counts$Drug, names(uv_results))]

# Combine all univariate tables into a single table after reordering
univariate_table <- tbl_stack(uv_results_ordered)

#bold the p's
univariate_table <- univariate_table %>%
  bold_p(t = 0.05)  

# Print the final univariate table
univariate_table

```


##Merge Tables

```{r}
# Merging the two tables with headers
merged_table <- tbl_merge(
  tbls = list(univariate_table, final_table),
  tab_spanner = c("**Univariate**", "**Multivariate**")
)

# Print the merged table
merged_table %>% as_gt() %>% gt::tab_options(table.font.names = "Arial")
```

##Export results
Write excel file with results output
```{r}
# Convert to huxtable and write to Excel
#merged_table %>%
 #as_hux_xlsx(file.path(wd$output,"RADIOHEADMultivariateTable.xlsx"))
```


#Multivariate with significant drugs in the same model

```{r} 
d1 <- dat %>%
  select(record_id, os, subject.dead.hospice, subject.age, prior.targeted_therapy, subject.metastatic.disease, Treatment, cancer_agreggate,stage, Opioid, Corticosteroid_systemic, Beta_Blockers, Diuretic, Laxative, Anti_emetic, Benzodiazepines)
         
# Relevel the covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")
```

```{r}
# Create the formula
formula <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                  paste(colnames(d1)[-(1:3)], collapse = "+"))

# Fit the Cox model
fit <- coxph(as.formula(formula), data = d1)

# Get the summary of the Cox model
summary_fit <- summary(fit)

# Print the results as a gtsummary table
gt_table <- tbl_regression(fit)
gt_table2 <-tbl_regression(fit, exponentiate = TRUE)
print(gt_table)
print(gt_table2)

```

```{r}
# Convert to huxtable and write to Excel
#gt_table %>%
 #as_hux_xlsx(file.path(wd$output,"combinedmultivariate_nopoly.xlsx"))

#gt_table2 %>%
# as_hux_xlsx(file.path(wd$output,"combinedmultivariate2_nopoly.xlsx"))
```

forest plot
```{r}
#convert model output to a tibble format using broom package
output <- tidy(fit, conf.int = TRUE)
#View(output)

#filter to only include variables we want visualized
c_forest <- output %>%
  filter(term %in% c("Opioid", "Corticosteroid_systemic", "Beta_Blockers", "Diuretic", "Laxative", "Anti_emetic", "Benzodiazepines"))
View(c_forest)
```

#visualize output via forest plot for figure 2 
```{r} 
#change names for visualization
c_forest <- c_forest %>%
  mutate(term = factor(case_when(
    term == "Opioid" ~ "Opioids",
    term == "Anti_emetic" ~ "Anti-emetics",
    term == "Diuretic" ~ "Diuretics", 
    term == "Corticosteroid_systemic" ~ "Corticosteroids (systemic)",
    term == "Beta_Blockers" ~ "Beta-blockers",
    term == "Laxative" ~ "Laxatives",
    TRUE ~ term))) %>%
  arrange(desc(estimate)) %>%
  mutate(term = factor(term, levels = rev(term))) #sort by ascending hazard ratios

#View(c_forest) 

```


```{r} 
forest_plot <- ggplot(c_forest, aes(x = estimate, y = term)) +
  geom_point(size = 3, color = "deepskyblue") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "deepskyblue") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
  scale_x_continuous(name = "Log Hazard Ratio", 
                     breaks = scales::pretty_breaks(), 
                     limits = c(-0.2, 1)) + # Set limits for x-axis
  scale_y_discrete(name = "") +
  theme_minimal(base_size = 22) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 25, color = "black"),
    axis.text.x = element_text(size = 25))

print(forest_plot)
    
    #add p-values
```
```{r}
#create p-value plot 
ps <- ggplot(c_forest, aes(x = 0, y = term,
            label = ifelse(p.value <0.001, "p < 0.001", sprintf("p = %.3f", p.value)))) +
   geom_text(
    aes(fontface = ifelse(p.value < 0.05, "bold", "plain")),  # Bold if p < 0.05
    hjust = 0, 
    size = 8
  ) +
  theme_void() +  # Removes background grid and axes
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(0, 0, 0, 10))  # Adjust right margin to give space for the labels
print(ps)

```

```{r}
#Combine plots 
# Define the layout 
layout <- c(
  area(t = 0, l = 0, b = 30, r = 14),  # Left plot (forest_plot), takes up the full height and 11 units of width
  area(t = 0, l = 12, b = 30, r = 18)   # Right plot (ps), positioned next to the left plot, with 2 units of width
)
f1 <- forest_plot + ps + plot_layout(design = layout)
print(f1)
```
```{r}
ggsave(file.path(wd$output, "Forest_plot_1updated.png"), plot = f1, width = 12, height = 6, dpi = 300)
```

##############################################################################################################
Check vifs: (to confirm these drug classes can be used in the same model)

```{r}
library(car)
# Create a linear model with the variables
model <- lm(Opioid ~ polypharmacy + Corticosteroid_systemic + Beta_Blockers + Diuretic + Laxative + Anti_emetic + Benzodiazepines + prior.targeted_therapy + stage + Treatment, data = sub_df)

# Calculate the VIF for this model
vif(model)

```

```{r}
# Check correlation between variables
cor(sub_df$Opioid, sub_df$polypharmacy)
```

########################################################################################

#Multivariate with opioids and palliative care interaction

first- create palliative care column 

```{r}
paldrugs <- c("Laxative", "Benzodiazepines", "Anti_emetic")
pall_df <- sub_df %>%
  mutate(palliative_drugs = ifelse(rowSums(across(all_of(paldrugs))) > 0, 1, 0))
write.csv(pall_df, "~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/palliativebundledf.csv", row.names = FALSE)
```
sanity check
```{r}
palcheck <- pall_df %>%
  select(record_id, all_of(paldrugs), palliative_drugs)
```
looks good

now multivariate with palliative drugs + opioids
```{r}
dat <- pall_df

d1 <- dat %>%
  select(record_id, os, subject.dead.hospice, subject.age, prior.targeted_therapy, subject.metastatic.disease,  Treatment, cancer_agreggate, stage,  Opioid, palliative_drugs)
         
# Relevel the covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")

```

```{r}
# Create the formula
formula <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                  paste(colnames(d1)[-(1:3)], collapse = "+"))

# Fit the Cox model
fit <- coxph(as.formula(formula), data = d1)

# Get the summary of the Cox model
summary_fit <- summary(fit)

# Print the results as a gtsummary table
gt_table <- tbl_regression(fit)
print(gt_table)
```


```{r}
# Convert to huxtable and write to Excel
#gt_table %>%
# as_hux_xlsx(file.path(wd$output,"Opioids_palliative_combined.xlsx"))
```

now with interaction term 

```{r}
# Create the formula with interaction term 
formula <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                  paste(colnames(d1)[-(1:3)], collapse = "+"),
                  " + Opioid:palliative_drugs")

# Fit the Cox model
fit <- coxph(as.formula(formula), data = d1)

# Get the summary of the Cox model
summary_fit <- summary(fit)

# Print the results as a gtsummary table
gt_table <- tbl_regression(fit)

#Calculate counts for each term 
opioid_count <- sum(d1$Opioid == 1, na.rm = TRUE)
palliative_count <- sum(d1$palliative_drugs == 1, na.rm = TRUE)
interaction_count <- sum(d1$Opioid == 1 & d1$palliative_drugs == 1, na.rm = TRUE)

interaction_counts <- c(
  Opioid = opioid_count,
  pallative_drugs = palliative_count,
  interaction = interaction_count) 
print(interaction_counts) 

gt_table2 <-tbl_regression(fit, exponentiate = TRUE)
#add these manually 

print(gt_table)
print(gt_table2)
```

```{r}
# Convert to huxtable and write to Excel
#gt_table %>%
 #as_hux_xlsx(file.path(wd$output,"Opioids_palliative_interaction.xlsx"))
# Convert to huxtable and write to Excel
#gt_table2 %>%
 #as_hux_xlsx(file.path(wd$output,"Opioids_palliative_interaction2.xlsx"))
```


forest plot
```{r}
#convert model output to a tibble format using broom package
output2 <- tidy(fit, conf.int = TRUE)
#View(output)

#filter to only include variables we want visualized
o_forest <- output2 %>%
  filter(term %in% c( "Opioid", "palliative_drugs", "Opioid:palliative_drugs"))
View(o_forest)
```

#visualize output via forest plot for figure 4
```{r} 
#change names for visualization
o_forest <- o_forest %>%
  mutate(term = factor(case_when(
    term == "Opioid" ~ "Opioids",
    term == "palliative_drugs" ~ "Palliative Bundle",
    term == "Opioid:palliative_drugs" ~ "Opioids * Palliative Bundle",
    TRUE ~ term))) %>%
  arrange(desc(estimate)) %>%
  mutate(term = factor(term, levels = rev(term)))#sort by ascending hazard ratios

View(o_forest) 

```


```{r} 
forest_plot2 <- ggplot(o_forest, aes(x = estimate, y = term)) +
  geom_point(size = 3, color = "orange") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "orange") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
  scale_x_continuous(name = "Log Hazard Ratio", 
                     breaks = scales::pretty_breaks(), 
                     limits = c(-0.5, 1)) + # Set limits for x-axis
  scale_y_discrete(name = "") +
  theme_minimal(base_size = 22) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 25, color = "black"),
    axis.text.x = element_text(size = 25))

print(forest_plot2)
```
add p-values
```{r}
#create p-value plot 
p2 <- ggplot(o_forest, aes(x = 0, y = term,
            label = ifelse(p.value <0.001, "p < 0.001", sprintf("p = %.3f", p.value)))) +
   geom_text(
    aes(fontface = ifelse(p.value < 0.05, "bold", "plain")),  # Bold if p < 0.05
    hjust = 0, 
    size = 8
  ) +
  theme_void() +  
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(0, 0, 0, 10))  # Adjust right margin to give space for the labels
print(p2)

```

```{r}
#Combine plots 
# Define the layout 
layout <- c(
  area(t = 0, l = 0, b = 20, r = 12),  # Left plot (forest_plot), takes up the full height and 11 units of width
  area(t = 0, l = 11, b = 20, r = 16)   # Right plot (ps), positioned next to the left plot, with 2 units of width
)
f2 <- forest_plot2 + p2 + plot_layout(design = layout)
print(f2)
```

```{r}
ggsave(file.path(wd$output, "Forest_plot_2.png"), plot = f2, width = 12, height = 6, dpi = 300)
```

## Individual drug multivariate
using individual_drugs from radiohead_dfs

```{r}
individual_drugs <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/individual_drugs_df.csv")
```

```{r}
colnames(individual_drugs)
```

```{r}
#first filter to drugs w/ at least 20 patients on them 
colsums2 <- colSums(individual_drugs[9:554])

sub2 <- names(colsums2[colsums2 >= 20])

indiv_drugs <- individual_drugs %>%
 select(1:10, 555:558, all_of(sub2))
```

```{r}
colnames(indiv_drugs)
```

```{r}
#remove supplements etc
i_drugs <- indiv_drugs %>%
  select(-"Iron", -"Folic_Acid", -"Multivitamin", -"Omega_3_Fatty_Acids", -"Vitamin_B12", -"Vitamin_C", -"Melatonin", -"Coenzyme_Q10", 
         -"Vitamin_B_Complex", -"No_meds", -"Potassium", -"Vitamin_D", -"Calcium", -"Probiotic", -"Magnesium", -"Glucosamine", -"Formoterol", -"Ipratropium", -"Albuterol", -"Vilanterol", -"Umeclidinium", -"Tiotropium")
colnames(i_drugs)
```

```{r}
d1 <- i_drugs
# Relevel the covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")

# List of baseline covariates
base_covariates <- c("subject.age", "prior.targeted_therapy", "subject.metastatic.disease", "Treatment", "cancer_agreggate", "stage")

# Initialize a list to store the results for each drug class
cox_results <- list()

# Loop through drug classes 
drug_columns <- names(d1)[14:92]

for (drug in drug_columns) {
  
   # Fit the Cox proportional hazards model with covariates and the drug class
  formula_str <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                        paste(base_covariates, collapse = " + "), " + ", drug)
  
  fit <- coxph(as.formula(formula_str), data = d1)
  
  # Summarize only the drug class in the table output (suppress the base covariates)
  tbl <- tbl_regression(fit, exponentiate = TRUE) %>%
    modify_table_body(~ .x %>%
      filter(variable == drug)) %>%
    modify_header(label = paste0("Drug: ", drug))
  
# Count the number of patients with a value of 1 in the covariate column
   count <- sum(d1[[drug]] == 1, na.rm = TRUE)
   
# Add the count as a new column in the table
  tbl <- tbl %>%
    modify_table_body(~ .x %>%
      mutate(Count = count)) %>%
    modify_header(Count = "**Count**")  # Add the count to the header
   
  # Store the results
  cox_results[[drug]] <- tbl
}

##Sorting by counts##
# Create a dataframe for sorting based on counts
drug_counts <- data.frame(Drug = drug_columns, Count = sapply(drug_columns, function(dc) sum(d1[[dc]] == 1, na.rm = TRUE)))

# Sort the drug counts in descending order
drug_counts <- drug_counts[order(-drug_counts$Count),]

# Reorder the cox_results based on sorted drug counts
cox_results <- cox_results[match(drug_counts$Drug, names(cox_results))] 

#Summarize the base covariates separately (fit a model without any drug class)#
base_formula_str <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                           paste(base_covariates, collapse = " + "))
base_fit <- coxph(as.formula(base_formula_str), data = d1)


# Combine the base covariates with the drug class tables
final_table <- tbl_stack((cox_results))

#bold the p's
final_table <- final_table %>%
  bold_p(t = 0.05)   %>%
   modify_header(label = "Drug Class")

# Print the final table
final_table
```


Add BH correction
```{r}
#Extract p-values 
p_values <- final_table$table_body$p.value 

#Perform BH correction
q_values <- p.adjust(p_values, method = "BH")

# Format q-values using gtsummary's style_pvalue()
q_values <- style_pvalue(q_values)

#Add to table 
final_table <- final_table %>% 
  modify_table_body(~ .x %>%
  mutate(`Q-value` = q_values)) %>%
  modify_header(`Q-value` = "**q-value**")

print(final_table)
  
```


```{r}
# Convert to huxtable and write to Excel
#final_table %>%
#as_hux_xlsx(file.path(wd$output,"individual_drugs_multivariate.xlsx"))
```



```{r}
beta_blockers <- c("Atenolol", "Bisoprolol", "Carvedilol", "Labetalol", "Metoprolol", 
                   "Nadolol", "Nebivolol", "Propranolol", "Sotalol") 

beta_df <- individual_drugs %>%
  select(1:8, "stage", "Treatment", "cancer_agreggate", all_of(beta_blockers))
```

```{r}
BB <- beta_df %>%
  mutate(cardioselective = case_when(
    Atenolol == 1 ~ 1,
    Bisoprolol == 1 ~ 1,
    Metoprolol == 1 ~ 1,
    Nebivolol == 1 ~ 1,
    TRUE ~ 0)) %>%
  mutate(noncardioselective = case_when(
    Carvedilol == 1 ~ 1,
    Labetalol == 1 ~ 1,
    Nadolol == 1 ~ 1,
    Propranolol == 1 ~ 1, 
    Sotalol == 1 ~ 1,
    TRUE ~ 0))
head(BB)

```
```{r}
table(BB$cardioselective, BB$noncardioselective)
```

```{r}
sum(BB$cardioselective) 
sum(BB$noncardioselective)
```


#Beta blocker multivariate 

```{r}
#Multivariate with significant drugs in the same model
dat <- BB
d1 <- dat %>%
  select(record_id, os, subject.dead.hospice, subject.age, prior.targeted_therapy, subject.metastatic.disease,  Treatment, cancer_agreggate, stage, cardioselective, noncardioselective)
         
# Relevel the covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")
```

```{r}
# Create the formula
formula <- paste0("Surv(os, subject.dead.hospice) ~ ", 
                  paste(colnames(d1)[-(1:3)], collapse = "+"))

# Fit the Cox model
fit <- coxph(as.formula(formula), data = d1)

# Get the summary of the Cox model
summary_fit <- summary(fit)

# Print the results as a gtsummary table
gt_table <- tbl_regression(fit)
gt_table2 <-tbl_regression(fit, exponentiate = TRUE)
print(gt_table)
print(gt_table2)
```
```{r}
# Convert to huxtable and write to Excel
#gt_table %>%
# as_hux_xlsx(file.path(wd$output,"betablockermultivariate.xlsx"))

#gt_table2 %>%
# as_hux_xlsx(file.path(wd$output,"betablockermultivariate2.xlsx"))
```

forest plot
```{r}
#convert model output to a tibble format using broom package
output <- tidy(fit, conf.int = TRUE)
#View(output)

#filter to only include variables we want visualized
b_forest <- output %>%
  filter(term %in% c("cardioselective", "noncardioselective"))
View(b_forest)
```

#visualize output via forest plot for figure 3
```{r} 
#change names for visualization
b_forest <- b_forest %>%
  mutate(term = factor(case_when(
    term == "cardioselective" ~ "Cardioselective Beta-Blockers",
    term == "noncardioselective" ~ "Non-Cardioselective Beta-Blockers",
    TRUE ~ term))) %>%
  arrange(desc(estimate)) %>%
  mutate(term = factor(term, levels = rev(term))) #sort by ascending hazard ratios

View(b_forest) 

```


```{r} 
forest_plot <- ggplot(b_forest, aes(x = estimate, y = term)) +
  geom_point(size = 4, color = "deepskyblue") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "deepskyblue") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
  scale_x_continuous(name = "Log Hazard Ratio", 
                     breaks = scales::pretty_breaks(), 
                     limits = c(-0.2, 1)) + # Set limits for x-axis
  scale_y_discrete(name = "") +
  theme_minimal(base_size = 20) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 20, color = "black"),
    axis.text.x = element_text(size = 20))

print(forest_plot)
    

```
 #add p-values
```{r}
#create p-value plot 
ps <- ggplot(b_forest, aes(x = 0, y = term,
            label = ifelse(p.value <0.001, "p < 0.001", sprintf("p = %.3f", p.value)))) +
   geom_text(
    aes(fontface = ifelse(p.value < 0.05, "bold", "plain")),  # Bold if p < 0.05
    hjust = 0, 
    size = 8
  ) +
  theme_void() +  # Removes background grid and axes
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(0, 0, 0, 10))  # Adjust right margin to give space for the labels
print(ps)

```

```{r}
#Combine plots 
# Define the layout 
layout <- c(
  area(t = 0, l = 0, b = 30, r = 16),  # Left plot (forest_plot), takes up the full height and 11 units of width
  area(t = 0, l = 12, b = 30, r = 20)   # Right plot (ps), positioned next to the left plot, with 2 units of width
)
f1 <- forest_plot + ps + plot_layout(design = layout)
print(f1)
```

```{r}
ggsave(file.path(wd$output, "Forest_plot_3.png"), plot = f1, width = 13, height = 6, dpi = 300)
```

#add univariate KMs
```{r}
drug_classes <- c("noncardioselective", "cardioselective") 

drug_labels <- c(
  "noncardioselective" = "Non-Cardioselective Beta-Blocker Exposure",
  "cardioselective" = "Cardioselective Beta-Blocker Exposure"
)
# Create survival object
surv_object <- Surv(time = BB$os, event = BB$subject.dead.hospice)   

# Create loop for plotting
for (drug in drug_classes) { 
  
  # Fit Kaplan-Meier curve
  fit <- survfit(surv_object ~ get(drug), data = BB)

  # Extract values
  cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ get(drug), data = BB)

  # Define custom colors for each drug class
  custom_colors <- c("orange", "deepskyblue")  # Adjust these hex codes to your preferred colors 
  
 # Dynamically set legend labels based on the drug name
  legend_labels <- c("No Exposure", drug_labels[drug])  # Uses the custom label from the mapping
  
  # Create the plot
  plot <- ggsurvplot(fit, data = BB,  pval.size = 9,  risk.table.fontsize = 8, pval = TRUE, conf.int = TRUE, 
                     risk.table = TRUE,
                     legend.labs = legend_labels,
                    palette = custom_colors, 
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.6, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 22), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 22), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
    plot.title = element_text(size = 22)
  ) +
  xlab("Time (months)")


  # Specify the output file path for each drug class
 # output_file <- file.path(wd$output, paste0("KM_", drug, ".png"))

  #Open a pdf device 
 # png(output_file, width = 8, height = 6, units = "in", res = 300)
  
  #print plot
  print(plot)
  
  #close device 
  #dev.off()
}
```
all beta blocker KM
```{r}
# Create survival object
surv_object <- Surv(time = sub_df$os, event = sub_df$subject.dead.hospice)   
  
  # Fit Kaplan-Meier curve
  fit <- survfit(surv_object ~ Beta_Blockers, data = sub_df)

  # Extract values
  cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ Beta_Blockers, data = sub_df)

  # Define custom colors for each drug class
  custom_colors <- c("orange", "deepskyblue")  # Adjust these hex codes to your preferred colors 
  
 # Dynamically set legend labels based on the drug name
  legend_labels <- c("No Beta-Blocker Exposure", "Beta-Blocker Exposure")  
  
  # Create the plot
  plot <- ggsurvplot(fit, data = sub_df, pval = TRUE,  pval.size = 9,  risk.table.fontsize = 8,  conf.int = TRUE, 
                     risk.table = TRUE,
                     legend.labs = legend_labels,
                    palette = custom_colors, 
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.6, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 22), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 22), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
    plot.title = element_text(size = 22)
  ) +
  xlab("Time (months)")
print(plot)

# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_Beta-Blockers.png")

# Open a PNG device
png(output_file_png, width = 10, height = 8, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()

```

Opioid KM
```{r}
# Create survival object
surv_object <- Surv(time = sub_df$os, event = sub_df$subject.dead.hospice)   
  
  # Fit Kaplan-Meier curve
  fit <- survfit(surv_object ~ Opioid, data = sub_df)

  # Extract values
  cox_mode <- coxph(Surv(os, subject.dead.hospice) ~ Opioid, data = sub_df)

  # Define custom colors for each drug class
  custom_colors <- c("orange", "deepskyblue")  # Adjust these hex codes to your preferred colors 
  
 # Dynamically set legend labels based on the drug name
  legend_labels <- c("No Opioid Exposure", "Opioid Exposure")  
  
  # Create the plot
  plot <- ggsurvplot(fit, data = sub_df, pval = TRUE,  pval.size = 9,  risk.table.fontsize = 8,  conf.int = TRUE, 
                     risk.table = TRUE,
                     legend.labs = legend_labels,
                    palette = custom_colors, 
                    legend.title = "")  # Set the custom color palette
  
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.6, 0.2),
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 22), # Axis titles
    legend.text = element_text(size = 25), # Legend labels
   legend.title = element_text(size = 22)
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 20),  # Axis tick labels
    axis.title = element_text(size = 22), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
    plot.title = element_text(size = 22)
  ) +
  xlab("Time (months)")
print(plot)

# Specify the output file path for PNG
output_file_png <- file.path(wd$output, "KMs/KM_Opioid.png")

# Open a PNG device
png(output_file_png, width = 10, height = 8, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()

```


