---
title: "04 - UMAP Visualization"
author: "Sammy Stone"
date: "2026-1-27"
output: html_document
---

# UMAP Visualization of Drug Class Exposure
#
# This script creates a UMAP projection of patients based on their drug class
# exposure patterns, highlighting minimal overlap between patients exposed to opioids and beta-blocker.
#
# Key Output:
# - Fig_2B.pdf (UMAP with opioid/beta-blocker overlap)
# - Umap_coordinates.csv (UMAP coordinates saved for reproducibility)
# - UMAP_model.rds (UMAP R object for reproducibility)
# - UMAP_parameters.csv (Umap parameters for reproducibility) 

# Important Info: 
# - Umap generated on 1-27-26 using umap: 0.2.10.0

# ==============================================================================
# Setup
# ==============================================================================
```{r setup, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(umap)

# Load project configuration
source(here::here("R", "config.R"))
```

# ==============================================================================
# Load Data
# ==============================================================================
```{r load-data}
# Load systemic drugs dataframe (for UMAP features)
systemic_df <- read.csv(paths$systemic_drugs)

# Load subclasses dataframe (for exposure categories)
cohortsub <- read.csv(paths$subclasses_df)

message(sprintf("✓ Loaded %d patients", nrow(systemic_df)))
```

# ==============================================================================
# Prepare UMAP Features
# ==============================================================================
```{r prepare-features}
# Filter and select features for UMAP
umap_features <- systemic_df %>%
  filter(!is.na(subject.dead.hospice)) %>%
  filter(stage != "UNK") %>%
  select(-c(meddra.disease.preferred.name, os, subject.dead.hospice, subject.age, line.1.therapy, subject.metastatic.disease, prior.targeted_therapy, irae.any, irae.severe, stage, Corticosteroids, record_id))  #Remove non medication columns or medications that are repeated


message(sprintf("✓ Using %d drug class features for UMAP", ncol(umap_features)))
message(sprintf("✓ Analyzing %d patients", nrow(umap_features)))
```

# ==============================================================================
# Run UMAP (with selected optimal parameters)
# ==============================================================================
```{r run-umap}
# Set seed for reproducibility
set.seed(123456)

# Configure UMAP with explicit parameters for reproducibility
umap_config <- umap.defaults
umap_config$random_state <- 123456
umap_config$n_neighbors <- 20      
umap_config$min_dist <- 0.3        
umap_config$spread <- 1.0          
umap_config$metric <- "euclidean"  
umap_config$n_epochs <- 200       

# Run UMAP
umap_result <- umap(umap_features, config = umap_config)

# Convert to dataframe 
umap_df <- data.frame(umap_result$layout)
colnames(umap_df) <- c("UMAP1", "UMAP2")

message("✓ UMAP projection complete")
message(sprintf("  Parameters: n_neighbors=%d, min_dist=%.1f, spread=%.1f",
               umap_config$n_neighbors, umap_config$min_dist, umap_config$spread))
```

# ==============================================================================
# Create Exposure Categories
# ==============================================================================
```{r create-exposure-categories}
# Filter cohortsub to match UMAP cohort
cohortsub_filtered <- cohortsub %>%
  filter(!is.na(subject.dead.hospice)) %>%
  filter(stage != "UNK")

# Create exposure categories
exposure_data <- cohortsub_filtered %>%
  select(record_id, Opioid, Beta_Blockers) %>%
  mutate(Exposure = case_when(
    Opioid == 1 & Beta_Blockers == 0 ~ "Opioids",
    Opioid == 0 & Beta_Blockers == 1 ~ "Beta-Blockers",
    Opioid == 1 & Beta_Blockers == 1 ~ "Overlap",
    TRUE ~ "Other"
  ))

# Add exposure to UMAP dataframe
umap_df$Exposure <- factor(
  exposure_data$Exposure,
  levels = c("Opioids", "Beta-Blockers", "Overlap", "Other")
)

# Summary of exposure categories
exposure_summary <- table(umap_df$Exposure)
message("\n=== Exposure Category Counts ===")
print(exposure_summary)
```

# ==============================================================================
# Save UMAP Coordinates 
# ==============================================================================
```{r save-umap-coordinates}
# Create data directory if it doesn't exist
umap_data_dir <- here::here("output", "data")
dir.create(umap_data_dir, showWarnings = FALSE, recursive = TRUE)

# Save UMAP coordinates with record IDs and exposure categories
umap_coordinates <- data.frame(
  record_id = cohortsub_filtered$record_id,
  UMAP1 = umap_df$UMAP1,
  UMAP2 = umap_df$UMAP2,
  Exposure = umap_df$Exposure
)

# Save as CSV
write.csv(
  umap_coordinates,
  here::here(umap_data_dir, "UMAP_coordinates.csv"),
  row.names = FALSE
)

# Also save the full UMAP object (includes model)
saveRDS(
  umap_result,
  here::here(umap_data_dir, "UMAP_model.rds")
)

# Save UMAP parameters for documentation
umap_params <- data.frame(
  parameter = c("seed", "random_state", "n_neighbors", "min_dist", "spread", "metric", "n_epochs"),
  value = c(123456, 
            umap_config$random_state,
            umap_config$n_neighbors,
            umap_config$min_dist,
            umap_config$spread,
            umap_config$metric,
            umap_config$n_epochs)
)

write.csv(
  umap_params,
  here::here(umap_data_dir, "UMAP_parameters.csv"),
  row.names = FALSE
)

```

# ==============================================================================
# Figure 2B: UMAP with Opioid/Beta-Blocker Overlap
# ==============================================================================
```{r figure-2b, fig.width=9, fig.height=7}
# Define custom colors
custom_colors <- c(
  "Opioids" = "#08306B",
  "Beta-Blockers" = "#FDAE61",
  "Overlap" = "#00BFFF",
  "Other" = "grey"
)

# Create UMAP plot
umap_plot <- ggplot(umap_df, aes(UMAP1, UMAP2, color = Exposure)) +
  geom_point(size = 1.5, alpha = 1) +
  scale_color_manual(values = custom_colors) +
  theme_minimal() +
  theme(
    legend.position = c(0.87, 0.85),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 20),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(color = "grey")
  ) +
  labs(x = "UMAP 1", y = "UMAP 2", color = NULL)

# Display plot
print(umap_plot)

# Save plot
ggsave(
  here::here("output", "figures", "Fig_2B.pdf"),
  plot = umap_plot,
  device = "pdf",
  width = 9,
  height = 7
)

```


# ==============================================================================
# Session Info
# ==============================================================================
```{r session-info}
message("\n=== Session Information ===")
sessionInfo()

message("\n=== Key Package Versions ===")
cat(sprintf("R version: %s\n", R.version.string))
cat(sprintf("umap: %s\n", packageVersion("umap")))
cat(sprintf("tidyverse: %s\n", packageVersion("tidyverse")))
cat(sprintf("ggplot2: %s\n", packageVersion("ggplot2")))
```
