---
title: "radiohead umap"
author: "Sammy Stone"
date: "2024-11-08"
output: html_document
---
```{r}
#using systemic_df

library(umap)
library(ggplot2)
library(tidyverse)
library(viridis)
library(uwot)
library(fastDummies)

```

Load:
```{r}
#systemic df
systemic_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_systemicdrugs.csv")

#sub_df
sub_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_multivariate_df.csv")

#cohortsub

cohortsub <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_subclasses.csv")

```


```{r}
df<- systemic_df
```

#medication UMAP

```{r}
top20 <- systemic_df %>%
  filter(!is.na(subject.dead.hospice)) %>%
  filter(stage != "UNK") %>%
  select(-c(2, 4, 5, 6, 7, 8, 11)) #remove cancer type, death event, age, line.1.therapy, prior.targeted therapy, mets, irae any, stage
```

any
```{r}

m.any <- top20 %>%
  select(-Corticosteroids, -irae.any, -irae.severe, -os, -record_id )
```

```{r}
colnames(m.any)
```

```{r}
#write_csv(m.any, "~/Desktop/RADIOHEAD_umap_df.csv")
```

```{r}
nrow(m.any)
```

```{r}
set.seed(123)
umap_config <- umap.defaults
umap_config$random_state <- 123

umap_result <- umap(m.any, n_neighbors = 20, min_dist = 0.3)
umap_df <- data.frame(umap_result)
colnames(umap_df) <- c("UMAP1", "UMAP2")
```

```{r}
# Ensure Opioid is a factor with levels 0 and 1
umap_df$Opioid <- factor(m.any$Opioid, levels = c(0, 1))

# Define custom colors
colors_1 <- c("0" = "deepskyblue", "1" = "orange")

# Plot
p <- ggplot(umap_df, aes(UMAP1, UMAP2, color = Opioid)) +
  geom_point() +
  scale_color_manual(values = colors_1) +  # Apply custom colors
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.8),
     axis.text = element_text(size = 18), 
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 22))  +
  labs(color = "Opioids") 

print(p)
```


```{r}
print(p)
```

```{r}
#ggsave("opioid_umap3.pdf", plot = p, device = "pdf", width = 7,  height = 7)
```

#UMAP Overlap
Fig 2B

```{r}
cohortsub2 <- cohortsub %>%
  filter(!is.na(subject.dead.hospice)) %>%
  filter(stage != "UNK")
```

```{r}
shared <- cohortsub2 %>%
  select(record_id, Opioid, Beta_Blockers) %>%
  mutate(Exposure = case_when(
    Opioid == 1 & Beta_Blockers == 0 ~ "Opioids",
    Opioid == 0 &  Beta_Blockers == 1 ~ "Beta-Blockers",
    Opioid == 1 & Beta_Blockers == 1 ~ "Overlap",
    TRUE ~ "Other"
  ))
```

```{r}
umap_df$Exposure <- factor(shared$Exposure, levels = c("Opioids", "Beta-Blockers", "Overlap", "Other"))

colors_2 <- c("Opioids" = "#08306B",
              "Beta-Blockers" = "#FDAE61",
              "Overlap" = "#00BFFF",
              "Other" = "grey")

```

```{r}
# Plot UMAP with multiple exposure categories
p <- ggplot(umap_df, aes(UMAP1, UMAP2, color = Exposure)) +
  geom_point(size = 1.5, alpha = 1) +  # Adjust transparency for better overlap visualization
  scale_color_manual(values = colors_2) +  
  theme_minimal() +  
  theme(
    legend.position = c(0.87, 0.85),
    axis.text = element_text(size = 18), 
    axis.title = element_text(size = 22),
    legend.text = element_text(size = 20),
    legend.title = element_blank(),
     panel.grid = element_blank(),  # Removes all grid lines
    panel.border = element_blank(), # Removes outer border
    axis.line = element_line(color = "grey")
  ) +
  labs(x = "UMAP 1", y = "UMAP 2", color = NULL)
print(p)
```
```{r}
ggsave("shared_umap.pdf", plot = p, device = "pdf", width = 9,  height = 7)
```



#Jaccards Index of Similarity- like overlap, but not calculated with respect to smaller size
calculate overlap between opioids with anti-emetic, laxative, diuretic, beta-blocker, systemic corticosteroids, benzodiazepines w/ percentages 

```{r}
# Define two sets
set1 <- c(1, 2, 3, 4, 5)
set2 <- c(3, 4, 5, 6, 7)

# Function to calculate Jaccard indec
calculate_jaccard<- function(set1, set2) {
  intersection <- intersect(set1, set2)
  union <- union(set1, set2)
  jaccard_coefficient <- length(intersection) /length(union)
  return(jaccard_coefficient)
}

# Calculate Overlap coefficient
jaccard <- calculate_jaccard(set1, set2)
print(jaccard)
```

define columns
```{r}
cols <- c("Anti_emetic", "Laxative", "Diuretic", "Beta_Blockers", "Corticosteroid_systemic", "Benzodiazepines")
```

loop 
```{r}
#initialize results list 
results <- list()

for (col in cols) {
  overlap <- calculate_jaccard(
    which(sub_df$Opioid == 1),
    which(sub_df[[col]] == 1)
  )
  results[[col]] <- overlap
} 

#convert results to a df
results_df <- data.frame(
  Drug_class = names(results), 
  Jaccard_Index = unlist(results)
)
results_df <- results_df %>%
  mutate(Drug_class = case_when(
    Drug_class == "Anti_emetic" ~ "Anti-emetics",
    Drug_class == "Laxative" ~ "Laxatives",
    Drug_class == "Beta_Blockers" ~ "Beta-Blockers",
    Drug_class == "Corticosteroid_systemic" ~ "Corticosteroid-systemic",
    Drug_class == "Diuretic" ~ "Diuretics",
   TRUE ~ Drug_class))
   
#save as excel 
#write.xlsx(results_df, "jaccardresults.xlsx")

View(results_df)

```

```{r}
results_df$Jaccard_Prop <- results_df$Jaccard_Index 

# Reorder the Drug_Class factor based on Percent_Overlap
results_df$Drug_Class <- factor(results_df$Drug_class, 
                                levels = results_df$Drug_class[order(results_df$Jaccard_Prop, decreasing = TRUE)])

# Create the bar plot with ordered Drug_Class and extended y-axis
ggplot(results_df, aes(x = Drug_Class, y = Jaccard_Prop)) +
  geom_bar(stat = "identity", fill = "deepskyblue", color = "grey", alpha=.6, width=.6) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "Jaccard Coefficient"
  ) +
   theme(
    axis.text = element_text(size = 14, color = "black"),  # Axis tick labels
    axis.title = element_text(size = 16, color = "black") # Axis titles
  ) +
  # Extend y-axis to 100%
  expand_limits(y = c(0, 1)) + 
  coord_flip() 
#geom_hline(yintercept = 25, linetype = "dashed", color = "darkgrey", linewidth = 0.5)  # Add horizontal dashed line at 50%

# Optionally save the plot
#ggsave("Jaccard_barplot.pdf")

```
manual check 
```{r}
opioid_emetic <- sub_df %>%
  filter(Opioid == 1 & Anti_emetic == 1)
nrow(opioid_emetic) 

emetic <- sub_df %>%
  filter(Anti_emetic == 1)
nrow(emetic)
```

#Beta-Blocker Jaccard 

define columns
```{r}
cols <- c("Anti_emetic", "Laxative", "Diuretic", "Opioid", "Corticosteroid_systemic", "Benzodiazepines")
```

loop 
```{r}
#initialize results list 
results <- list()

for (col in cols) {
  overlap <- calculate_jaccard(
    which(sub_df$Beta_Blockers == 1),
    which(sub_df[[col]] == 1)
  )
  results[[col]] <- overlap
} 

#convert results to a df
results_df <- data.frame(
  Drug_class = names(results), 
  Jaccard_Index = unlist(results)
)
results_df <- results_df %>%
  mutate(Drug_class = case_when(
    Drug_class == "Anti_emetic" ~ "Anti-emetics",
    Drug_class == "Laxative" ~ "Laxatives",
    Drug_class == "Opioid" ~ "Opioids",
    Drug_class == "Corticosteroid_systemic" ~ "Corticosteroid-systemic",
    Drug_class == "Diuretic" ~ "Diuretics",
   TRUE ~ Drug_class))
   
#save as excel 
#write.xlsx(results_df, "jaccardresults.xlsx")

View(results_df)

```

```{r}
results_df$Jaccard_Prop <- results_df$Jaccard_Index 

# Reorder the Drug_Class factor based on Percent_Overlap
results_df$Drug_Class <- factor(results_df$Drug_class, 
                                levels = results_df$Drug_class[order(results_df$Jaccard_Prop, decreasing = TRUE)])

# Create the bar plot with ordered Drug_Class and extended y-axis
ggplot(results_df, aes(x = Drug_Class, y = Jaccard_Prop)) +
  geom_bar(stat = "identity", fill = "deepskyblue", color = "grey", alpha=.6, width=.6) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "Jaccard Coefficient"
  ) +
   theme(
    axis.text = element_text(size = 14, color = "black"),  # Axis tick labels
    axis.title = element_text(size = 16, color = "black") # Axis titles
  ) +
  # Extend y-axis to 100%
  expand_limits(y = c(0, 1)) + 
  coord_flip() 
#geom_hline(yintercept = 25, linetype = "dashed", color = "darkgrey", linewidth = 0.5)  # Add horizontal dashed line at 50%

# Optionally save the plot
#ggsave("BB_Jaccard_barplot.pdf")

```
manual checks
```{r}
opioid_BB <- sub_df %>%
  filter(Opioid == 1 & Beta_Blockers == 1)
nrow(opioid_BB) 

opioid_check <- sub_df %>%
  filter(Opioid == 1)
  nrow(opioid_check)

opioid_check2 <- sub_df %>%
  filter(Opioid == 1 & Beta_Blockers == 0)
nrow(opioid_check2)

bb_check <- sub_df %>%
  filter(Beta_Blockers == 1)
nrow(bb_check)

bb_check2 <- sub_df %>%
  filter(Beta_Blockers == 1 & Opioid == 1)
nrow(bb_check2)
```

