---
title: "Opioid KMs"
author: "Sammy Stone"
date: "2024-12-04"
output: html_document
---
#load

```{r}
library(tidyverse)
library(readr)
library(openxlsx)
library(ggvenn)
library(survminer)
library(survival)
library(eulerr)
```
```{r}
#set directories
wd <- list()
wd$main <- file.path("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD")
wd$data <- file.path(wd$main, "data")
wd$dataP <- file.path(wd$data, "processed")
wd$output <- file.path(wd$main, "output_sammy")
wd$script <- file.path(wd$main, "scripts_sammy")
wd$opioid <- file.path(wd$output, "opioid KMs")
```

```{r}
pall_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/palliativebundledf.csv")

sub_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_multivariate_df.csv")
```

#(50+) w/ sub classes of interest and stage and polypharmacy

using sub_df from Radiohead_dfs.Rmd (file with polypharmacy, and drugs + sub classes of interest with at least 50 patients)

```{r}
dat <- sub_df %>%
  select(-"X", -"Bronchodilator")
```


#Opioid Survival Under Different Conditions
Goal- interrogate how baseline opioid use influences survival outcomes using KM curves 


#Stage IV

```{r}
stage4 <- sub_df %>%
  filter(stage == "iv")
```


```{r}
km_model <- survfit(Surv(os, subject.dead.hospice) ~ Opioid, data = stage4)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(km_model, data = stage4, pval = TRUE, pval.size = 7,  risk.table.fontsize = 7, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("No Opioid Exposure", "Opioid Exposure"),
                   palette = custom_colors,
                   legend.title = "") 

# Apply theme_minimal with additional text modifications
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 18), # Legend labels
   legend.title = element_text(size = 20), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 18), # Legend labels
    plot.title = element_text(size = 20),
  ) +
  xlab("Time (months)")
print(plot)
```

```{r}

# Specify the output file path for each drug class
output_file_png <- file.path(wd$output, "Opioidstage4.png")

# Open a PNG device
png(output_file_png, width = 8, height = 7, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()


```


#Opioids stages 0-III
with 0 mets
```{r}
zeromet <- sub_df %>%
 filter(subject.metastatic.disease == 0 & !stage == "iv") 
```

```{r}
km_model <- survfit(Surv(os, subject.dead.hospice) ~ Opioid, data = zeromet)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(km_model, data = zeromet, pval = TRUE, pval.size = 7,  risk.table.fontsize = 7, conf.int = FALSE, 
                   risk.table = TRUE,
                   legend.labs = c("No Opioid Exposure", "Opioid Exposure"),
                   palette = custom_colors,
                   legend.title = "") 

# Apply theme_minimal with additional text modifications
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
   legend.title = element_text(size = 20), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
    plot.title = element_text(size = 20),
  ) +
  xlab("Time (months)")

print(plot)

# Specify the output file path for each drug class
output_file_png <- file.path(wd$opioid, "Opioids_nomets2.png")

# Open a PNG device
png(output_file_png, width = 8, height = 7, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()

```


Compare this to the palliative bundle (using pall_df) 


#Palliative drugs only
stages 0-3 no mets

```{r}
earlystage2 <- pall_df %>%
  filter(stage %in% c("0-ii", "iii") & subject.metastatic.disease == 0)
```


```{r}
km_model <- survfit(Surv(os, subject.dead.hospice) ~ palliative_drugs, data = earlystage2)

# Define custom colors for each drug class
custom_colors <- c("orange", "deepskyblue")

# Create the plot
plot <- ggsurvplot(km_model, data = earlystage2,  pval = TRUE, pval.size = 7,  risk.table.fontsize = 7,  conf.int = FALSE, 
                   risk.table = TRUE,
                   legend.labs = c("No Palliative Bundle Exposure", "Palliative Bundle Exposure"),
                   legend.title = "",
                   palette = custom_colors) 

# Apply theme_minimal with additional text modifications
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.70, 0.2),
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
   legend.title = element_text(size = 20), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 20), # Legend labels
    plot.title = element_text(size = 20),
  ) +
  xlab("Time (months)")
print(plot)
```


```{r}
# Specify the output file path for each drug class
output_file_png <- file.path(wd$opioid, "PB_stages0to3nomet.png")

# Open a PNG device
png(output_file_png, width = 8, height = 7, units = "in", res = 300)

# Print the plot
print(plot)

# Close the PNG device
dev.off()

```


#Venn 
```{r}
head(pall_df)
```
subset for convenience
```{r}
pall <- pall_df %>%
  select(record_id, Opioid, palliative_drugs)
```

```{r}
# Count categories
only_opioid <- sum(pall$Opioid == 1 & pall$palliative_drugs == 0)
only_palliative <- sum(pall$Opioid == 0 & pall$palliative_drugs == 1)
both <- sum(pall$Opioid == 1 & pall$palliative_drugs == 1)

# Total counts including the overlap
total_opioids <- only_opioid + both    
total_palliative <- only_palliative + both 

# Euler diagram using non-disjoint counts
fit <- euler(c(
  "Opioids" = total_opioids,
  "Palliative Bundle" = total_palliative,
  "Opioids&Palliative Bundle" = both
))

```

```{r}
# Plot with transparent fill and colored outlines
plot(fit,
     fills = NA,  # No fill color
     edges = list(col = c("skyblue", "orange"), lwd = 2),  # Colored outlines
     labels = list(font = 2),
     quantities = TRUE,
     main = "")
```
```{r}
# Euler diagram using *disjoint* counts
fit <- euler(c(
  "Opioids" = only_opioid,
  "Palliative Bundle" = only_palliative,
  "Opioids&Palliative Bundle" = both
))

par(family = "Arial") 

# Plot with transparent fill and colored outlines
p <- plot(fit,
     fills = NA,  # No fill color
     edges = list(col = c("skyblue", "orange"), lwd = 2),  # Colored outlines
     labels = list(font = 2),
     quantities = TRUE,
     main = "")
p
```

```{r}
# Specify the output file path for each drug class
output_file_png <- file.path(wd$output, "venn.png")

# Open a PNG device
png(output_file_png, width = 8, height = 6, units = "in", res = 300)

# Print the plot
p

# Close the PNG device
dev.off()
```

#Stacked KM
```{r}
stack_df <- pall_df %>%
  select(-X) %>%
  select(record_id, os, subject.dead.hospice, stage, subject.metastatic.disease, Opioid, palliative_drugs) #pall drugs includs laxitive, antiemetic, benzo, opioid not specificed

#new cats 
stacked_df <- stack_df %>%
  mutate(category = case_when(
    Opioid == 1 & palliative_drugs == 0 ~ "Opioids Only",
    Opioid == 0 & palliative_drugs == 1 ~ "Palliative Only",
    Opioid == 1 & palliative_drugs == 1 ~ "Opioids and Palliative Drugs",
    Opioid == 0 & palliative_drugs == 0 ~ "No Opioids or Palliative Drugs"))

stacked_df <- stacked_df %>%
  mutate(category = factor(category,
                              levels = c("No Opioids or Palliative Drugs", "Palliative Only", "Opioids Only", "Opioids and Palliative Drugs")))
```

```{r}
km_model <- survfit(Surv(os, subject.dead.hospice) ~ category, data = stacked_df)

# Define custom colors for each drug class
custom_colors <- c( "grey", "orange", "deepskyblue", "blue")

# Create the plot
plot <- ggsurvplot(km_model, data = stacked_df, pval = FALSE, pval.size = 7,  risk.table.fontsize = 7, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("No Opioids or Palliative Drugs", "Palliative Only", "Opioids Only", "Opioids and Palliative Drugs"),
                   palette = custom_colors,
                   legend.title = "") 

# Apply theme_minimal with additional text modifications
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 18), # Legend labels
   legend.title = element_text(size = 20), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 18), # Legend labels
    plot.title = element_text(size = 20),
  ) +
  xlab("Time (months)")
print(plot)
```

```{r}
# Get pairwise comparisons with multiple comparison correction
pairwise_test <- pairwise_survdiff(Surv(os, subject.dead.hospice) ~ category, 
                                  data = stacked_df,
                                  p.adjust.method = "holm")  # or "bonferroni", "BH", "fdr"

print(pairwise_test)
```

#Stacked KM early stage

```{r}
early_stacked <- stacked_df %>%
 filter(subject.metastatic.disease == 0 & !stage == "iv") 
```

```{r}
km_model <- survfit(Surv(os, subject.dead.hospice) ~ category, data = early_stacked)

# Define custom colors for each drug class
custom_colors <- c( "grey", "orange", "deepskyblue", "blue")

# Create the plot
plot <- ggsurvplot(km_model, data = early_stacked, pval = TRUE, pval.size = 7,  risk.table.fontsize = 7, conf.int = TRUE, 
                   risk.table = TRUE,
                   legend.labs = c("No Opioids or Palliative Drugs", "Palliative Only", "Opioids Only", "Opioids and Palliative Drugs"),
                   palette = custom_colors,
                   legend.title = "") 

# Apply theme_minimal with additional text modifications
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 18), # Legend labels
   legend.title = element_text(size = 20), 
  ) +
  xlab("Time (months)")

# Apply theme_minimal for the risk table 
plot$table <- plot$table + 
  theme_minimal()  +
  theme(
    axis.text = element_text(size = 18),  # Axis tick labels
    axis.title = element_text(size = 20), # Axis titles
    legend.text = element_text(size = 18), # Legend labels
    plot.title = element_text(size = 20),
  ) +
  xlab("Time (months)")
print(plot)
```

```{r}
# Get pairwise comparisons with multiple comparison correction
pairwise_test <- pairwise_survdiff(Surv(os, subject.dead.hospice) ~ category, 
                                  data = early_stacked,
                                  p.adjust.method = "holm")  # or "bonferroni", "BH", "fdr"

print(pairwise_test)
```
```

