---
title: "05 - Opioid and Palliative Drug Survival Analysis"
author: "Sammy Stone"
date: "2026-1-27"
output: html_document
---

# Opioid and Palliative Drug Survival Analysis
#
# This script performs Kaplan-Meier survival analysis to assess the association
# between opioid and palliative drug exposure and overall survival in early-stage
# cancer patients (stages 0-III with no metastases).
#
# Key Analyses:
# 1. Opioid exposure KM in stages 0-III (no metastases)
# 2. Palliative bundle exposure KM in stages 0-III (no metastases)
# 3. Venn diagram of opioid and palliative bundle overlap
#
# Outputs:
# - Figures: Fig_3A.png, Fig_3C.png,
#            Fig_3D.png

# ==============================================================================
# Setup
# ==============================================================================
```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(survminer)
library(survival)
library(eulerr)
library(here)
```

```{r load-config}
# Load project configuration
source(here::here("R", "config.R"))
```
```{r load-data}
# Load processed dataframes
pall_df <- read.csv(palliative_bundle_df)
sub_df <- read.csv(multivariate_df)

message("palliative bundle dataframe: ", nrow(pall_df), " patients")
message(" multivariate dataframe: ", nrow(sub_df), " patients")
```

# ==============================================================================
# Opioid Survival in Early Stage Disease (Stages 0-III, No Metastases)
# ==============================================================================
**Fig 3C**
```{r opioid-early-stage}
# Filter for stages 0-III with no metastatic disease
zeromet <- sub_df %>%
  filter(subject.metastatic.disease == 0 & !stage == "iv")

message(sprintf("\nFiltered to stages 0-III, no metastases: %d patients", nrow(zeromet)))
message(sprintf("  Opioid exposure: %d patients", sum(zeromet$Opioid == 1)))
message(sprintf("  No opioid exposure: %d patients", sum(zeromet$Opioid == 0)))
```
```{r opioid-km-plot}
# Fit Kaplan-Meier model
km_model <- survfit(Surv(os, subject.dead.hospice) ~ Opioid, data = zeromet)

# Define custom colors
custom_colors <- c("orange", "deepskyblue")

# Create the survival plot
plot <- ggsurvplot(
  km_model, 
  data = zeromet, 
  pval = TRUE, 
  pval.size = 7,  
  risk.table.fontsize = 7, 
  conf.int = FALSE, 
  risk.table = TRUE,
  legend.labs = c("No Opioid Exposure", "Opioid Exposure"),
  palette = custom_colors,
  legend.title = ""
)

# Apply theme and formatting to main plot
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20)
  ) +
  xlab("Time (months)")

# Apply theme to risk table
plot$table <- plot$table + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    plot.title = element_text(size = 20)
  ) +
  xlab("Time (months)")

print(plot)
```
```{r save-opioid-plot}
# Save the plot
output_file_png <- here::here(figures, "Fig_3C.png")
png(output_file_png, width = 8, height = 7, units = "in", res = 300)
print(plot)
dev.off()

message("✓ Saved Fig_3C.png")
```

# ==============================================================================
# Palliative Bundle Survival in Early Stage Disease (Stages 0-III, No Metastases)
# ==============================================================================
**Fig 3D**
```{r palliative-early-stage}
# Filter for stages 0-III with no metastatic disease
earlystage2 <- pall_df %>%
  filter(stage %in% c("0-ii", "iii") & subject.metastatic.disease == 0)

message(sprintf("\nFiltered to stages 0-III, no metastases: %d patients", nrow(earlystage2)))
message(sprintf("  Palliative bundle exposure: %d patients", sum(earlystage2$palliative_drugs == 1)))
message(sprintf("  No palliative bundle exposure: %d patients", sum(earlystage2$palliative_drugs == 0)))
```
```{r palliative-km-plot}
# Fit Kaplan-Meier model
km_model <- survfit(Surv(os, subject.dead.hospice) ~ palliative_drugs, data = earlystage2)

# Define custom colors
custom_colors <- c("orange", "deepskyblue")

# Create the survival plot
plot <- ggsurvplot(
  km_model, 
  data = earlystage2,  
  pval = TRUE, 
  pval.size = 7,  
  risk.table.fontsize = 7,  
  conf.int = FALSE, 
  risk.table = TRUE,
  legend.labs = c("No Palliative Bundle Exposure", "Palliative Bundle Exposure"),
  legend.title = "",
  palette = custom_colors
)

# Apply theme and formatting to main plot
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.70, 0.2),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20)
  ) +
  xlab("Time (months)")

# Apply theme to risk table
plot$table <- plot$table + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    plot.title = element_text(size = 20)
  ) +
  xlab("Time (months)")

print(plot)
```
```{r save-palliative-plot}
# Save the plot
output_file_png <- here::here(figures, "Fig_3D.png")
png(output_file_png, width = 8, height = 7, units = "in", res = 300)
print(plot)
dev.off()

message("✓ Saved Fig_3D.png")
```

# ==============================================================================
# Overlap Between Opioid and Palliative Bundle Exposure
# ==============================================================================
```{r venn-data-prep}
# Select relevant columns
pall <- pall_df %>%
  select(record_id, Opioid, palliative_drugs)

# Count categories
only_opioid <- sum(pall$Opioid == 1 & pall$palliative_drugs == 0)
only_palliative <- sum(pall$Opioid == 0 & pall$palliative_drugs == 1)
both <- sum(pall$Opioid == 1 & pall$palliative_drugs == 1)

message("\nOverlap between opioids and palliative bundle:")
message(sprintf("  Opioids only: %d patients", only_opioid))
message(sprintf("  Palliative bundle only: %d patients", only_palliative))
message(sprintf("  Both: %d patients", both))
message(sprintf("  Neither: %d patients", nrow(pall) - only_opioid - only_palliative - both))
```
```{r euler-diagram}
# Create Euler diagram using disjoint counts
fit <- euler(c(
  "Opioids" = only_opioid,
  "Palliative Bundle" = only_palliative,
  "Opioids&Palliative Bundle" = both
))

# Set font family
par(family = "Arial") 

# Plot with transparent fill and colored outlines
p <- plot(
  fit,
  fills = NA,
  edges = list(col = c("skyblue", "orange"), lwd = 2),
  labels = list(font = 2),
  quantities = TRUE,
  main = ""
)

p
```
```{r save-venn}
# Save the Venn diagram
output_file_png <- here::here(figures, "Fig_3A.png")
png(output_file_png, width = 8, height = 6, units = "in", res = 300)
plot(
  fit,
  fills = NA,
  edges = list(col = c("skyblue", "orange"), lwd = 2),
  labels = list(font = 2),
  quantities = TRUE,
  main = ""
)
dev.off()

message("✓ Saved Fig_3A.png")
```

# ==============================================================================
# Session Info
# ==============================================================================
```{r session-info}
sessionInfo()
```
