---
title: "06 - Demographics"
author: "Sammy Stone"
date: "2024-09-30"
output: html_document
---

# Demographic Comparisons
#
# This script generates demographic summary tables for the RADIOHEAD cohort,
# comparing patient characteristics by opioid and beta-blocker exposure.
#
# Key Analyses:
# 1. Demographic summary by opioid exposure (with chi-square tests)
# 2. Demographic summary by beta-blocker exposure (with chi-square tests)
#
# Outputs:
# - Tables: Table_S3.xlsx (opioid demographics), Table_S5.xlsx (beta-blocker demographics)

# ==============================================================================
# Setup
# ==============================================================================
```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(gtsummary)
library(here)
library(cardx)
```

```{r load-config}
# Load project configuration
source(here::here("R", "config.R"))
```

```{r load-data}
# Load processed dataframes
sub_df <- read.csv(multivariate_df)

dem_df <- read.csv(demographic_df)

message("✓ Loaded multivariate dataframe: ", nrow(sub_df), " patients")
message("✓ Loaded demographic dataframe: ", nrow(dem_df), " patients")
```

```{r merge-sex}
# Extract sex column from demographic dataframe
sex_col <- dem_df %>%
  select(record_id, subject.sex)

# Merge sex column with multivariate dataframe
dem <- left_join(sub_df, sex_col, by = "record_id")

message("✓ Merged sex column into analysis dataframe")
```

# ==============================================================================
# Opioid Exposure Demographics
# ==============================================================================
```{r opioid-demographics}
# Create factor labels for opioid exposure
dem_opioid <- dem %>%
  mutate(Opioid = factor(Opioid,
                         levels = c(0, 1),
                         labels = c("No Opioid Exposure", "Opioid Exposure")))

message("\nOpioid exposure distribution:")
message(sprintf("  No opioid exposure: %d patients", sum(dem$Opioid == 0)))
message(sprintf("  Opioid exposure: %d patients", sum(dem$Opioid == 1)))

# Create summary table with statistical tests
opioid_tbl <- dem_opioid %>%
  tbl_summary(
    include = c(subject.age, subject.sex, cancer_agreggate, stage, polypharmacy, 
                prior.targeted_therapy, subject.metastatic.disease, Treatment),
    by = Opioid, 
    missing = "no",
    label = list(
      subject.age ~ "Patient Age",
      cancer_agreggate ~ "Cancer",
      polypharmacy ~ "Cumulative Drug Class Exposures",
      stage ~ "Stage",
      prior.targeted_therapy ~ "Prior Targeted Therapy (Y/N)",
      subject.metastatic.disease ~ "Metastatic Disease (Y/N)",
      subject.sex ~ "Patient Sex"
    )
  ) %>%
  add_p() %>%                          
  bold_p(t = 0.05) %>%                  
  add_q(method = "bonferroni")

# Print the table
print(opioid_tbl)
```

```{r save-opioid-table}
# Save as Excel file
opioid_tbl %>%
  as_hux_xlsx(here::here(tables, "Table_S3.xlsx"))

message("✓ Saved Table_S3.xlsx (opioid demographics)")
```

# ==============================================================================
# Beta-Blocker Exposure Demographics
# ==============================================================================
```{r bb-demographics}
# Create factor labels for beta-blocker exposure
dem_bb <- dem %>%
  mutate(Beta_Blockers = factor(Beta_Blockers,
                                levels = c(0, 1),
                                labels = c("No Beta Blocker Exposure", "Beta Blocker Exposure")))

message("\nBeta-blocker exposure distribution:")
message(sprintf("  No beta-blocker exposure: %d patients", sum(dem$Beta_Blockers == 0)))
message(sprintf("  Beta-blocker exposure: %d patients", sum(dem$Beta_Blockers == 1)))

# Create summary table with statistical tests
bb_tbl <- dem_bb %>%
  tbl_summary(
    include = c(subject.age, subject.sex, cancer_agreggate, stage, polypharmacy, 
                prior.targeted_therapy, subject.metastatic.disease, Treatment),
    by = Beta_Blockers, 
    missing = "no",
    label = list(
      subject.age ~ "Patient Age",
      cancer_agreggate ~ "Cancer",
      polypharmacy ~ "Polypharmacy (continuous)",
      stage ~ "Stage",
      prior.targeted_therapy ~ "Prior Targeted Therapy (Y/N)",
      subject.metastatic.disease ~ "Metastatic Disease (Y/N)",
      subject.sex ~ "Patient Sex"
    )
  ) %>%
  add_p() %>%                          
  bold_p(t = 0.05) %>%                  
  add_q(method = "bonferroni")

# Print the table
print(bb_tbl)
```

```{r save-bb-table}
# Save as Excel file
bb_tbl %>%
  as_hux_xlsx(here::here(tables, "Table_S5.xlsx"))

message("✓ Saved Table_S5.xlsx (beta-blocker demographics)")
```

# ==============================================================================
# Summary
# ==============================================================================
```{r summary}
message("\n=== Demographic Analysis Complete ===")
message("\nOutput Files:")
message("Tables:")
message("  - output/tables/Table_S3.xlsx (opioid demographics)")
message("  - output/tables/Table_S5.xlsx (beta-blocker demographics)")
```

# ==============================================================================
# Session Info (for reproducibility)
# ==============================================================================
```{r session-info}
sessionInfo()
```
