---
title: "RADIOHEAD meds demographics"
author: "Sammy Stone"
date: "2024-09-30"
output: html_document
---

Goal: Generate summaries of radiohead demographics for entire cohort + sub-cohorts of interest, then perform chi square test to assess differences 

Data: using the dem_df from radiohead_dfs.Rmd using sub_df with sex column pulled from dem_df (new- 05/14/25)

#Pre-processing
Load packages
```{r}
library(tidyverse)
library(readr)
library(writexl)
library(tableone)
library(gtsummary)
```

Establish working directories 

```{r}
wd <- list()
wd$main <- file.path("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD")
wd$data <- file.path(wd$main, "data")
wd$dataP <- file.path(wd$data, "processed")
wd$output <- file.path(wd$main, "output_sammy")
wd$script <- file.path(wd$main, "scripts")
```

Load data
```{r}
sub_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_multivariate_df.csv")
sub_df <- sub_df %>%
  select(-X, "Bronchodilator")
dem_df <- read.csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_demograpphic_df.csv")
dem_df <- dem_df %>%
  select(-X)
```

Update cancer column to match the multivariable analysis ## ignore in update
```{r}
dem_df <- dem_df %>%
  mutate(Cancer = ifelse(meddra.disease.preferred.name %in% 
  c("Non-small cell lung cancer", "Malignant melanoma", "Renal cell carcinoma", "Bladder cancer", "Small cell lung cancer"),
  meddra.disease.preferred.name,
  "Other")) 
```


##May 2025 Update 

```{r}
head(dem_df)

sex_col <- dem_df %>%
  select(record_id, subject.sex) 

head(sex_col)
nrow(sex_col)
```

create updated dem df by merging sex column from old dem_df with sub_df
```{r}
dem <- left_join(sub_df, sex_col, by = "record_id")

View(dem)
```

#JUNE 2025 Update
#pan-cancer vs opioid-- restrict opioids vs cohort without opioids


```{r}

dem <- dem %>%
  mutate(Opioid = factor(Opioid,
                                levels = c(0, 1),
                                labels = c("No Opioid Exposure", "Opioid Exposure")))
# Create a summary table
comparison_tbl <- dem %>%
 tbl_summary(
    include = c(subject.age, subject.sex, cancer_agreggate, stage, polypharmacy, prior.targeted_therapy, subject.metastatic.disease, Treatment),
    by = Opioid, 
    missing = "no",
    label = list(subject.age ~ "Patient Age",
                 cancer_agreggate ~ "Cancer",
                 polypharmacy ~ "Cumulative Drug Class Exposures",
                 stage ~ "Stage",
                 prior.targeted_therapy ~"Prior Targeted Therapy (Y/N)",
                 subject.metastatic.disease ~"Metastatic Disease (Y/N)",
                 subject.sex ~ "Patient Sex")
  ) %>%
  add_p() %>%                          
  bold_p(t = 0.05) %>%                  
  add_q(method = "bonferroni")         

# Print the table
comparison_tbl
    
```

```{r}
# Convert to huxtable and write to Excel
comparison_tbl %>%
 as_hux_xlsx(file.path(wd$output,"/Demographic Comparisons/opioid_demographics2.xlsx"))
```

#BB's
```{r}

dem <- dem %>%
  mutate(Beta_Blockers = factor(Beta_Blockers,
                                levels = c(0, 1),
                                labels = c("No Beta Blocker Exposure", "Beta Blocker Exposure")))

# Create a summary table
comparison_tbl <- dem %>%
 tbl_summary(
    include = c(subject.age, subject.sex, cancer_agreggate, stage, polypharmacy, prior.targeted_therapy, subject.metastatic.disease, Treatment),
    by = Beta_Blockers, 
    missing = "no",
    label = list(subject.age ~ "Patient Age",
                 cancer_agreggate ~ "Cancer",
                 polypharmacy ~ "Polypharmacy (continuous)",
                 stage ~ "Stage",
                 prior.targeted_therapy ~"Prior Targeted Therapy (Y/N)",
                 subject.metastatic.disease ~"Metastatic Disease (Y/N)",
                 subject.sex ~ "Patient Sex")
  ) %>%
  add_p() %>%                          
  bold_p(t = 0.05) %>%                  
  add_q(method = "bonferroni")         

# Print the table
comparison_tbl
    
```


```{r}
# Convert to huxtable and write to Excel
comparison_tbl %>%
 as_hux_xlsx(file.path(wd$output,"bb_demographics2.xlsx"))
```

