---
title: "07 - Supplementary Tables"
author: "Sammy Stone"
date: "2025-01-30"
output: html_document
---

# Supplementary Table S1: Drug Classification Summary
#
# This script generates a comprehensive summary of all medications prescribed
# to patients in the RADIOHEAD cohort, organized by drug class, subclass, and
# individual drug name. It includes patient counts and prescription frequencies.
#
# Outputs:
# - Tables: Table_S1.xlsx (comprehensive drug classification summary)

# ==============================================================================
# Setup
# ==============================================================================
```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(openxlsx)
library(readxl)
library(here)
```
```{r load-config}
# Load project configuration
source(here::here("R", "config.R"))
```
```{r load-data}
# Load processed dataframes
sub_df <- read_csv(multivariate_df)

# Load per-patient harmonized medication data
perpatient <- read_excel(here::here(data_raw, "Per_patient_harmonized_round_10.xlsx"))

message("✓ Loaded multivariate dataframe: ", nrow(sub_df), " patients")
message("✓ Loaded per-patient medication data: ", nrow(perpatient), " medication entries")
```

# ==============================================================================
# Filter and Clean Medication Data
# ==============================================================================
```{r filter-medications}
# Filter out non-systemic medications (vitamins, supplements, drop_na)
perpatient_filt <- perpatient %>%
  mutate(Revised_Classification = str_trim(Revised_Classification)) %>%
  filter(!Revised_Classification %in% c("Drop_NA", "NA", "Vitamin", "Supplement", "Supplement-non systemic"))

message("\nFiltered medications:")
message(sprintf("  Original entries: %d", nrow(perpatient)))
message(sprintf("  After filtering: %d", nrow(perpatient_filt)))
```
```{r filter-cohort}
# Select relevant columns
perpatient_filt2 <- perpatient_filt %>%
  select(record_id, harmonized_final, Revised_Classification)

# Get record IDs of analyzed cohort (1024 patients)
records <- sub_df %>%
  select(record_id)

# Filter to match 1024 patient cohort
perpatient_filt3 <- perpatient_filt2 %>%
  filter(record_id %in% records$record_id)

message(sprintf("\nFiltered to analyzed cohort: %d patients", length(unique(perpatient_filt3$record_id))))

# Check for missing patients
missing_records <- records %>%
  filter(!record_id %in% perpatient_filt2$record_id)

if (nrow(missing_records) > 0) {
  message(sprintf("  Note: %d patients only on vitamins/supplements (excluded)", nrow(missing_records)))
}
```
```{r consolidate-corticosteroids}
# Consolidate inhaled corticosteroids into non-systemic category
perpatient_filt4 <- perpatient_filt3 %>%
  mutate(Revised_Classification = ifelse(Revised_Classification == "Corticosteroid-inhaled", 
                                         "Corticosteroid-non systemic", 
                                         Revised_Classification))

message(sprintf("\nFinal medication dataset: %d entries", nrow(perpatient_filt4)))
```

# ==============================================================================
# Define Drug Subclasses
# ==============================================================================
```{r define-subclasses}
# Mental health medications
benzodiazepines <- c("Clonazepam", "Alprazolam", "Lorazepam", "Diazepam", 
                     "Temazepam", "Chlordiazepoxide", "Clobazam")

ssris <- c("Escitalopram", "Citalopram", "Fluoxetine", "Paroxetine", "Sertraline")

# Cardiovascular medications
beta_blockers <- c("Metoprolol", "Carvedilol", "Atenolol", "Propranolol", "Sotalol", 
                   "Nebivolol", "Nadolol", "Labetalol", "Bisoprolol")

ACEi_ARB <- c("Lisinopril", "Enalapril", "Benazepril", "Ramipril", "Quinapril",
              "Fosinopril", "Perindopril", "Trandolapril", "Losartan", "Valsartan", 
              "Irbesartan", "Olmesartan", "Candesartan", "Azilsartan", "Telmisartan")

cc_blockers <- c("Amlodipine", "Diltiazem", "Nifedipine", "Felodipine", "Verapamil")

# Cholesterol lowering
statins <- c("Atorvastatin", "Simvastatin", "Rosuvastatin", "Pravastatin", 
             "Lovastatin", "Fluvastatin")

# Anti-pyretic/anti-inflammatory
nsaid <- c("NSAID", "Ibuprofen", "Naproxen", "Meloxicam", "Diclofenac", "Celecoxib", 
           "Indomethacin", "Ketorolac", "Nabumetone", "Nepafenac")

message("\nDefined drug subclasses:")
message(sprintf("  Beta-blockers: %d drugs", length(beta_blockers)))
message(sprintf("  ACEi/ARBs: %d drugs", length(ACEi_ARB)))
message(sprintf("  Calcium channel blockers: %d drugs", length(cc_blockers)))
message(sprintf("  Statins: %d drugs", length(statins)))
message(sprintf("  Benzodiazepines: %d drugs", length(benzodiazepines)))
message(sprintf("  SSRIs: %d drugs", length(ssris)))
message(sprintf("  NSAIDs: %d drugs", length(nsaid)))
```
```{r create-subclass-df}
# Create subclass mapping dataframe
beta_blocker_df <- data.frame(Drug_Name = beta_blockers, Drug_SubClass = "Beta_Blockers")
ACEi_ARB_df <- data.frame(Drug_Name = ACEi_ARB, Drug_SubClass = "ACEi_ARB")
cc_blocker_df <- data.frame(Drug_Name = cc_blockers, Drug_SubClass = "CC_blockers")
statins_df <- data.frame(Drug_Name = statins, Drug_SubClass = "Statin")
benzo_df <- data.frame(Drug_Name = benzodiazepines, Drug_SubClass = "Benzodiazepines")
NSAID_df <- data.frame(Drug_Name = nsaid, Drug_SubClass = "NSAIDs")
ssri_df <- data.frame(Drug_Name = ssris, Drug_SubClass = "SSRIs")

# Combine all subclass mappings
subclass_df <- rbind(beta_blocker_df, ACEi_ARB_df, cc_blocker_df, statins_df, 
                     benzo_df, NSAID_df, ssri_df)
```
```{r merge-subclasses}
# Rename columns and merge with subclass assignments
perpatient_4 <- perpatient_filt4 %>%
  rename(Drug_Name = harmonized_final, 
         Drug_Class = Revised_Classification)

# Left join to add subclass information
perptsub <- left_join(perpatient_4, subclass_df, by = "Drug_Name")

# Select final columns
perpatient.sub <- perptsub %>%
  select(record_id, Drug_Name, Drug_SubClass, Drug_Class)

message(sprintf("\nMerged subclass assignments: %d entries", nrow(perpatient.sub)))
```

# ==============================================================================
# Generate Summary Tables
# ==============================================================================
```{r summarize-individual-drugs}
# Summary at individual drug level
perpt_sum <- perpatient.sub %>%
  group_by(Drug_Class, Drug_SubClass, Drug_Name) %>%
  summarize(
    Patient_Count = n_distinct(record_id),
    Cumulative_Count = n(),
    .groups = "drop"
  ) %>%
  mutate(Patient_Freq = (Patient_Count / 1024) * 100)

message("\nSummary at drug level:")
message(sprintf("  Unique drugs: %d", nrow(perpt_sum)))
```
```{r summarize-subclasses}
# Summary at drug subclass level
perpt_sum2 <- perpatient.sub %>%
  group_by(Drug_Class, Drug_SubClass) %>%
  summarize(
    Patient_Count = n_distinct(record_id),
    Cumulative_Count = n(),
    .groups = "drop"
  ) %>%
  mutate(Patient_Freq = (Patient_Count / 1024) * 100)

message(sprintf("  Unique subclasses: %d", sum(!is.na(perpt_sum2$Drug_SubClass))))
```
```{r summarize-classes}
# Summary at drug class level
perpt_sum3 <- perpatient.sub %>%
  group_by(Drug_Class) %>%
  summarize(
    Patient_Count = n_distinct(record_id),
    Cumulative_Count = n(),
    .groups = "drop"
  ) %>%
  mutate(Patient_Freq = (Patient_Count / 1024) * 100)

message(sprintf("  Unique drug classes: %d", nrow(perpt_sum3)))
message(sprintf("  Total medication entries: %d", sum(perpt_sum3$Cumulative_Count)))
```
```{r combine-summaries}
# Combine all summary levels
combined_sum <- bind_rows(perpt_sum3, perpt_sum2, perpt_sum)

# Sort by drug class and subclass
combined_sorted <- combined_sum %>%
  select(Drug_Class, Drug_SubClass, Drug_Name, Patient_Count, Patient_Freq, Cumulative_Count) %>%
  arrange(Drug_Class, Drug_SubClass)

message("\nCombined summary table:")
message(sprintf("  Total rows: %d", nrow(combined_sorted)))
message(sprintf("  Drug classes represented: %d", n_distinct(combined_sorted$Drug_Class)))
```

# ==============================================================================
# Save Output
# ==============================================================================
```{r save-table}
# Save as Excel file (Table S1)
write.xlsx(combined_sorted, here::here(tables, "Table_S1.xlsx"))

message("\n✓ Saved Table_S1.xlsx")
```

# ==============================================================================
# Summary Statistics
# ==============================================================================
```{r summary-stats}
message("\n=== Supplementary Table S1 Complete ===")
message("\nKey Statistics:")
message(sprintf("  Total patients analyzed: 1024"))
message(sprintf("  Total medication entries: %d", sum(perpt_sum3$Cumulative_Count)))
message(sprintf("  Unique drug classes: %d", n_distinct(combined_sorted$Drug_Class)))
message(sprintf("  Unique drug subclasses: %d", sum(!is.na(unique(combined_sorted$Drug_SubClass)))))
message(sprintf("  Unique individual drugs: %d", sum(!is.na(unique(combined_sorted$Drug_Name)))))

message("\nOutput Files:")
message("Tables:")
message("  - output/tables/Table_S1.xlsx")
```

# ==============================================================================
# Session Info (for reproducibility)
# ==============================================================================
```{r session-info}
sessionInfo()
```

##83 Drug classes, 82 when you exclude NA, 81 excluding Drop-NA, and 80 excluding cancer. Clean up performed in excel.
