---
title: "TableS1"
author: "Sammy Stone"
date: "2025-01-30"
output: html_document
---
```{r}
library(tidyverse)
library(openxlsx)
library(readr)
```


```{r}
wd <- list()
wd$main <- "/Users/samanthastone/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD"
wd$data <- file.path(wd$main, "data")
wd$dataP <- file.path(wd$data, "processed")
wd$output <- file.path(wd$main, "output_sammy")
wd$script <- file.path(wd$main, "scripts_sammy_final")
```


```{r}
sub_df <- read_csv("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/data/important_dataframes_sammy/RADIOHEAD_multivariate_df.csv")
```

##Use the per-patient harmonized round 10 file

1. add sub classes 
2. filter out supplements/vitamins/drop.na 
3. filter out patient ids not on 
2. group by record id, then get counts for each individual drug + revised classification 
3. without grouping get total counts

```{r}
perpatient <- read_excel("~/Library/CloudStorage/Box-Box/ZW_015_RADIOHEAD/output/Per_patient_harmonized_round_10.xlsx")
```

```{r}
perpatient_filt <- perpatient %>%
  mutate(Revised_Classification = str_trim(Revised_Classification)) %>%
  filter(!Revised_Classification %in% c("Drop_NA","NA", "Vitamin", "Supplement", "Supplement-non systemic"))
#not filtering, may need to manually remove after
```

```{r}
table(perpatient$Revised_Classification)
```

```{r}
perpatient_filt2 <- perpatient_filt %>%
  select(record_id, harmonized_final, Revised_Classification) 
```

```{r}
#get record ids of 1024 analysed pts
records <- sub_df %>%
  select(record_id) 
View(records)
```


```{r}
#filter patients to match 1024 cohort 
perpatient_filt3 <- perpatient_filt2 %>%
  filter(record_id %in% records$record_id) 
#View(perpatient_filt3)
print(length(unique(perpatient_filt3$record_id)))
```
```{r}
missing_records <- records %>%
  filter(!record_id %in% perpatient_filt2$record_id)

print(nrow(missing_records))  # Number of missing records
print(missing_records)  # See which record IDs are missing
```
missing patients are all on vitamins/supplements 

```{r}
perpatient_filt4 <- perpatient_filt3 %>%
  mutate(Revised_Classification = ifelse(Revised_Classification == "Corticosteroid-inhaled", "Corticosteroid-non systemic", Revised_Classification))
```

```{r}
table(perpatient_filt3$Revised_Classification)

```

```{r}
table(perpatient_filt4$Revised_Classification)
```

#create subclass df 


```{r}
#mental health
benzodiazepines <- c("Clonazepam", "Alprazolam", "Lorazepam", "Diazepam", "Temazepam", "Chlordiazepoxide", "Clobazam")

ssris <- c("Escitalopram", "Citalopram", "Fluoxetine", "Paroxetine", "Sertraline")

#cardiovascular

beta_blockers <- c("Metoprolol", "Carvedilol", "Atenolol", "Propranolol", "Sotalol", 
                   "Nebivolol", "Nadolol", "Labetalol", "Bisoprolol")

ACEi_ARB <- c("Lisinopril", "Enalapril", "Benazepril", "Ramipril", "Quinapril",
          "Fosinopril", "Perindopril", "Trandolapril", "Losartan", "Valsartan", "Irbesartan", "Olmesartan", "Candesartan", "Azilsartan","Telmisartan")

cc_blockers <- c("Amlodipine", "Diltiazem", "Nifedipine", "Felodipine", "Verapamil")
  
#Cholesterol Lowing
statins <- c("Atorvastatin","Simvastatin", "Rosuvastatin", "Pravastatin", "Lovastatin", "Fluvastatin")


#Anti-pyretic

nsaid <- c("NSAID", "Ibuprofen", "Naproxen", "Meloxicam", "Diclofenac", "Celecoxib", "Indomethacin", 
           "Ketorolac", "Nabumetone", "Nepafenac")

```



```{r}
# Create a data frame for each list and assign the drug class
beta_blocker_df <- data.frame(Drug_Name = beta_blockers, Drug_SubClass = "Beta_Blockers")
ACEi_ARB_df <- data.frame(Drug_Name = ACEi_ARB, Drug_SubClass = "ACEi_ARB")
cc_blocker_df <- data.frame(Drug_Name = cc_blockers, Drug_SubClass = "CC_blockers") 
statins_df <- data.frame(Drug_Name = statins, Drug_SubClass = "Statin") 
benzo_df <- data.frame(Drug_Name = benzodiazepines, Drug_SubClass = "Benzodiazepines")
NSAID_df <- data.frame(Drug_Name = nsaid, Drug_SubClass = "NSAIDs")
ssri_df <- data.frame(Drug_Name = ssris, Drug_SubClass = "SSRIs")

# Combine all data frames into one
subclass_df <- rbind(beta_blocker_df, ACEi_ARB_df, cc_blocker_df, statins_df, benzo_df, NSAID_df, ssri_df)

#View(subclass_df)
```

```{r}
perpatient_4 <- perpatient_filt4 %>%
 rename(Drug_Name = harmonized_final, 
        Drug_Class = Revised_Classification)
```


```{r}
perptsub <- left_join(perpatient_4, subclass_df, by = "Drug_Name") 
View(perptsub)
```

```{r}
perpatient.sub <- perptsub %>%
  select(record_id, Drug_Name, Drug_SubClass, Drug_Class)
```

summary table (summarize drug class, sub class, and individual drug)

```{r}
#summarize at drug level
perpt_sum <- perpatient.sub %>%
  group_by(Drug_Class, Drug_SubClass, Drug_Name) %>% #group each distinct combo
  summarize(Patient_Count = n_distinct(record_id), #get counts per patient
           Cumulative_Count = n(),
           .groups = "drop") %>%
    mutate(Patient_Freq = (Patient_Count/1024)*100) 
#View(perpt_sum)
  #  mutate(Cumulative_Freq = (Patient_Count/(sum(Cumulative_Count))) *100)

#print(sum(perpt_sum$Cumulative_Count))
#7695
```

summarize just drug class and drug sub class
```{r}
perpt_sum2 <- perpatient.sub %>%
  group_by(Drug_Class, Drug_SubClass) %>% #group by drug/subclass combo
  summarize(Patient_Count = n_distinct(record_id), #get counts per patient
           Cumulative_Count = n(),
           .groups = "drop") %>%
    mutate(Patient_Freq = (Patient_Count/1024)*100)

  #View(perpt_sum2.1)
```

summarize just drug class 

```{r}
perpt_sum3 <- perpatient.sub %>%
  group_by(Drug_Class) %>% #group by drug/subclass combo
  summarize(Patient_Count = n_distinct(record_id), #get counts per patient
           Cumulative_Count = n(),
           .groups = "drop") %>%
    mutate(Patient_Freq = (Patient_Count/1024)*100)

  #View(perpt_sum3)
```

```{r}
combined_sum <- bind_rows(perpt_sum3, perpt_sum2, perpt_sum) 

combined_sorted <- combined_sum %>%
  select(Drug_Class, Drug_SubClass, Drug_Name, Patient_Count, Patient_Freq, Cumulative_Count) %>%
  arrange(Drug_Class, Drug_SubClass)

View(combined_sorted)
```

```{r}
#write.xlsx(combined_sorted, "~/Desktop/FigS1.xlsx")
```
#sanity check
```{r}
View(combined_sorted)
```

manually remove duplicates with drug class

get cumulative total count
```{r}
print(sum(perpt_sum3$Cumulative_Count))
```
```{r}
perpt_sum5 <- perpatient.sub %>%
  group_by(Drug_Name) %>% #group by drug/subclass combo
  summarize(Patient_Count = n_distinct(record_id), #get counts per patient
           Cumulative_Count = n(),
           .groups = "drop") %>%
    mutate(Patient_Freq = (Patient_Count/1024)*100)

View(perpt_sum5)
```

```{r}
print(sum(perpt_sum5$Cumulative_Count))
```

7695 individual drugs prescribed 

```{r}
# Assuming your dataframe is named df
num_unique_classes <- combined_sorted %>% 
  summarise(unique_count = n_distinct(Drug_Class))

print(num_unique_classes) 

length(unique(combined_sorted$Drug_Class))
```


```{r}
unique_classes <- unique(combined_sorted$Drug_Class)
print(unique_classes)
```
83 Drug classes, 82 when you exclude NA, 81 excluding Drop-NA, and 80 excluding cancer
