---
title: "08_additional_analyses"
author: "Sammy Stone"
date: "2026-01-27"
output: html_document
---

# Additional Analyses
#
# This script generates reviewer tables and performs analyses performed during the revision period.
#
# Key Analyses:
# 1. rwPFS survival analysis (multivariable output similar to Table 1, but using rwPFS instead of rwOS)
# 2. Sub-cohort (single cancer) beta-blocker and opioid survival analysis (univariable and multivariable top 5 cancer types + other)
# 3.Beta-blocker vs Cardiac co-morbidity analyses
#
# Outputs:
# - Tables S2, S4, S5 and S6
#-  Figures: S1, Reviewer Table/Figure 1

# ==============================================================================
# Setup
# ==============================================================================
```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(gtsummary)
library(here)
library(openxlsx)
library(readxl)
library(survival)
library(survminer)
library(viridis)
library(huxtable)
```

```{r load-config}
# Load project configuration
source(here::here("R", "config.R"))
```

```{r load-data}
# Load processed dataframes
sub_df <- read.csv(multivariate_df)
message("multivariate dataframe: ", nrow(sub_df), " patients") 

cardiac <- openxlsx::read.xlsx(cardiac)
message("Cardiac Conditions dataframe: ", nrow(cardiac), " patients") 

df_clean <- read.csv(df_clean)
message("df clean dataframe: ", nrow(df_clean), " patients") 

cards <- openxlsx::read.xlsx(cards)

```


# ==============================================================================
# Summary of Cancer Type x Beta-Blocker/Opioid Exposure
# ==============================================================================

get a sense of breadth of cancer types/counts for opioids/beta blockers

```{r}
# Create new dataframe with just cancer type, opioid + beta-blocker columns
df1 <- sub_df %>%
  select(meddra.disease.preferred.name, Opioid, Beta_Blockers)
```

```{r}
# Tabulate patient counts for each cancer type and by drug exposure
df2 <- df1 %>% 
  group_by(meddra.disease.preferred.name) %>%
  summarize(
    patients = n(), 
    opioid_n = sum(Opioid == 1), 
    beta_blocker_n = sum(Beta_Blockers == 1), 
    .groups = "drop"
  ) %>%
  arrange(desc(patients))
```


```{r}
# Refine so that patient count is included in medication columns
df3 <- df2 %>% 
  mutate(
    Opioids = paste0(opioid_n, "/", patients),
    `Beta-Blockers` = paste0(beta_blocker_n, "/", patients)) %>%
  rename(Cancer = meddra.disease.preferred.name) %>%
  select(Cancer, Opioids, `Beta-Blockers` )
```

```{r}
# Save
  write.xlsx(df3, here::here(tables, "Reviewer_Table_1.xlsx"))
```
# ==============================================================================
# Progression Free Survival (rwPFS) Analyses
# ==============================================================================

```{r}
# Integrate PFS with sub_df
sub_df <- read.csv(paths$multivariate_df)

# get pfs
pfs <- df_clean %>%
  select(subject.id, probable.pfs, probable.pfs.event) %>%
  rename(record_id = subject.id) 

# truncate pfs
pfs <- pfs %>%
  mutate(probable.pfs.truncated = ifelse(probable.pfs >= 15, 15, probable.pfs)) 

# remove old def
pfs <- pfs %>%
  select(-probable.pfs) %>%
  rename(probable.pfs = probable.pfs.truncated) 

# join 
sub_df <- left_join(sub_df, pfs, by = "record_id")

# rename 
dat <- sub_df

# verify no NA patients and numeric structure
d1 <- dat %>% 
  filter(!is.na(subject.dead.hospice))

d1 <- dat %>% 
  filter(!is.na(probable.pfs.event))


```

Step 1: Multivariate Cox Regression (All Drug Classes)

**Table S2**
```{r}
# Relevel the covariates
d1$Treatment <- relevel(factor(d1$Treatment), ref = "pd1-pdl1")
d1$cancer_agreggate <- relevel(factor(d1$cancer_agreggate), ref = "Non-small cell lung cancer")
d1$stage <- relevel(factor(d1$stage), ref = "iv")

# List of baseline covariates
base_covariates <- c("subject.age", "prior.targeted_therapy", "subject.metastatic.disease", "Treatment", "cancer_agreggate", "stage")

# Initialize a list to store the results for each drug class
cox_results <- list()

# Loop through drug classes 
drug_class_columns <- names(d1)[11:41]

for (drug_class in drug_class_columns) {
  
   # Fit the Cox proportional hazards model with covariates and the drug class
  formula_str <- paste0("Surv(probable.pfs, probable.pfs.event) ~ ", 
                        paste(base_covariates, collapse = " + "), " + ", drug_class)
  
  fit <- coxph(as.formula(formula_str), data = d1)
  
  # Create summary table (show only drug class, suppress covariates)
  tbl <- tbl_regression(fit, exponentiate = TRUE) %>%
    modify_table_body(~ .x %>%
      filter(variable == drug_class)) %>%
    modify_header(label = paste0("Drug Class: ", drug_class))
  
# Add patient count
   count <- sum(d1[[drug_class]] == 1, na.rm = TRUE)
   
# Add count as column
  tbl <- tbl %>%
    modify_table_body(~ .x %>%
      mutate(Count = count)) %>%
    modify_header(Count = "**Count**")  # Add the count to the header
   
  # Store the results
  cox_results[[drug_class]] <- tbl
}


# Sort by count
drug_counts <- data.frame(Drug = drug_class_columns,
                          Count = sapply(drug_class_columns, function(dc) sum(d1[[dc]] == 1, na.rm = TRUE)))

# Sort the drug counts in descending order
drug_counts <- drug_counts[order(-drug_counts$Count),]

# Reorder the cox_results based on sorted drug counts
cox_results <- cox_results[match(drug_counts$Drug, names(cox_results))] 

# Create a stacked gtsummary table
final_table <- tbl_stack(cox_results) %>%
   modify_header(label = "Drug Class")

#bold the p's
final_table <- final_table %>%
  bold_p(t = 0.05)  

# Print the final table
final_table 

```



Step 2: Add Benjamini-Hochberg Correction
```{r bh-correction}
# Extract p-values
p_values <- final_table$table_body$p.value

# Perform BH correction
q_values <- p.adjust(p_values, method = "BH")
q_values <- style_pvalue(q_values)

# Add q-values to table
final_table <- final_table %>%
  modify_table_body(~ .x %>% mutate(`Q-value` = q_values)) %>%
  modify_header(`Q-value` = "**q-value**")

print(final_table)
```

Step 3: Univariate Cox Regression (All Drug Classes)
```{r}
# Initialize a list 
uv_results <- list()

# Loop through each drug class 
for (drug_class in drug_class_columns) {
  
  # Fit a univariable Cox model
  formula_uv <- as.formula(paste0("Surv(probable.pfs, probable.pfs.event) ~ `", drug_class, "`"))
  fit_uv <- coxph(formula_uv, data = d1)
  
  # Create a table summarizing the results 
  tbl_uv <- tbl_regression(fit_uv, exponentiate = TRUE) %>%
     modify_header(label = "Covariate") %>%
    modify_table_body(~ .x %>% mutate(variable = drug_class)) 
  
  # Store results
  uv_results[[drug_class]] <- tbl_uv
}
# Use the drug counts to reorder
uv_results_ordered <- uv_results[match(drug_counts$Drug, names(uv_results))]

# Combine all tables into a single table after reordering
univariate_table <- tbl_stack(uv_results_ordered)

#bold the p's
univariate_table <- univariate_table %>%
  bold_p(t = 0.05)  

# Print the final table
univariate_table
```


Step 4: Merge Univariate and Multivariate Tables
```{r merge-tables}
# Merge tables with headers
merged_table <- tbl_merge(
  tbls = list(univariate_table, final_table),
  tab_spanner = c("**Univariate**", "**Multivariate**")
)

# Display as formatted table
merged_table %>% 
  as_gt() %>% 
  gt::tab_options(table.font.names = "Arial")

# Save merged table to Excel
merged_table %>%
  as_hux_xlsx(here::here("output", "tables", "Table_S2_PFS.xlsx"))

```

# ==============================================================================
# Single Cancer type opioid + beta-blocker KMs
# ==============================================================================
** Tables S5 and S6 **
```{r}
# Create aggregated treatment column 
single_df <- sub_df %>%
  mutate(
    treatment.three = case_when(
    sub_df$Treatment == "pd1-pdl1" ~ "pd1-pdl1",
    sub_df$Treatment == "chemo+ICI"  ~ "ICI + other",
    sub_df$Treatment == "ctla4+pd1-pdl1" ~ "ctla4+pd1-pdl1",
    sub_df$Treatment == "other+pd1-pdl1" ~ "ICI + other",
    TRUE ~ NA_character_)) %>%
  mutate(treatment.three = as.factor(treatment.three)) 


```


Running as a function to make more efficient, covariate adjustment depends on cancer type (ie sclc too small for prior
targeted therapy), all cancers except other use treatment.three

```{r}
# Create function that performs univariable and multivariable analyses
create_combined_tables <- function(data,
                                   drug_var,
                                   time_var = "os",
                                   event_var = "subject.dead.hospice",
                                   base_covariates = c("subject.age", 
                                                      "prior.targeted_therapy", 
                                                      "subject.metastatic.disease", 
                                                      "stage",
                                                      "treatment.three"),
                                   cancer_type = c("Non-small cell lung cancer", 
                                                  "Malignant melanoma", 
                                                  "Renal cell carcinoma", 
                                                  "Bladder cancer",
                                                  "Small cell lung cancer",
                                                  "Other"),
                                   output_dir = here::here("output", "tables")) {

  
  # Initialize results lists
  combined_results <- list()
  model_objects <- list()  # to check later
  
  # Loop through each cancer type
  for (cancer in cancer_type) {
    
    # Filter data for this cancer type
    cancer_data <- data %>% 
      filter(cancer_agreggate == cancer) %>%
      filter(!is.na(.data[[event_var]]))
    
    message(paste("Processing", cancer, "- N =", nrow(cancer_data)))
    
    # Skip if no data (shouldn't be the case but sanity step)
    if (nrow(cancer_data) == 0) {
      message(paste("  ✗ Skipping - no data"))
      next
    }
    
    # Calculate drug exposure counts/proportion
    n_total <- nrow(cancer_data)
    n_exposed <- sum(cancer_data[[drug_var]] == 1, na.rm = TRUE)
    exposure_ratio <- paste0(n_exposed, "/", n_total)
    
    # --- UNIVARIABLE MODEL ---
    formula_uv <- as.formula(paste0("Surv(", time_var, ", ", event_var, ") ~ `", drug_var, "`"))
    
    uv_result <- tryCatch({
      fit_uv <- coxph(formula_uv, data = cancer_data)
      tbl_uv <- tbl_regression(fit_uv, exponentiate = TRUE) %>%
        bold_p(t = 0.05)
      list(success = TRUE, table = tbl_uv)
    }, error = function(e) {
      message(paste("  ✗ Univariate error:", e$message))
      list(success = FALSE, table = NULL)
    })
    
    # --- MULTIVARIABLE MODEL ---
    # Relevel stage if it exists
    if ("stage" %in% names(cancer_data)) {
      cancer_data$stage <- relevel(factor(cancer_data$stage), ref = "iv")
    }
    
    # Relevel treatment variables if they exist
    if ("Treatment" %in% names(cancer_data)) {
      cancer_data$Treatment <- relevel(factor(cancer_data$Treatment), ref = "pd1-pdl1")
    }
    if ("treatment.three" %in% names(cancer_data)) {
      cancer_data$treatment.three <- relevel(factor(cancer_data$treatment.three), ref = "pd1-pdl1")
    }
    
    # Modify covariates based on cancer type
    modified_covariates <- base_covariates
    
    # For "Other" cancer type, use original Treatment instead of treatment.three
    if (cancer == "Other") {
      modified_covariates <- gsub("treatment.three", "Treatment", modified_covariates)
      message("  Note: Using  original 'Treatment' for Other cancer type")
    }
    
    # For SCLC, exclude prior.targeted_therapy
    if (cancer == "Small cell lung cancer") {
      modified_covariates <- modified_covariates[modified_covariates != "prior.targeted_therapy"]
      message("  Note: Excluded 'prior.targeted_therapy' for SCLC due to sparse data")
    }
    
    # Check which covariates exist
    available_covariates <- modified_covariates[modified_covariates %in% names(cancer_data)]
    
    mv_result <- if (length(available_covariates) > 0) {
      formula_str <- paste0("Surv(", time_var, ", ", event_var, ") ~ ", 
                           paste(available_covariates, collapse = " + "), 
                           " + `", drug_var, "`")
      
      tryCatch({
        fit_mv <- coxph(as.formula(formula_str), data = cancer_data)
        
        # Save the model object for inspection
        model_objects[[cancer]] <- fit_mv
        
        tbl_mv <- tbl_regression(fit_mv, exponentiate = TRUE) %>%
          modify_table_body(~ .x %>% filter(variable == drug_var)) %>%
          bold_p(t = 0.05)
        list(success = TRUE, table = tbl_mv)
      }, error = function(e) {
        message(paste("  ✗ Multivariable error:", e$message))
        list(success = FALSE, table = NULL)
      })
    } else {
      message(paste("  ✗ No covariates available for multivariable model"))
      list(success = FALSE, table = NULL)
    }
    
    # --- MERGE TABLES ---
    if (uv_result$success && mv_result$success) {
      
      # Merge univariate and multivariable
      combined_tbl <- tbl_merge(
        tbls = list(uv_result$table, mv_result$table),
        tab_spanner = c("**Univariable**", "**Multivariable**")
      ) %>%
        modify_header(label = "**Cancer Type**") %>%
        modify_table_body(~ .x %>% mutate(
          label = cancer,
          n_exposed_total = exposure_ratio
        )) %>%
        modify_header(n_exposed_total = "**N Exposed/Total**") %>%
        modify_table_body(~ .x %>% relocate(n_exposed_total, .after = label))
      
      combined_results[[cancer]] <- combined_tbl
      message(paste("  ✓ Combined analysis complete -", exposure_ratio, "exposed"))
      
    } else if (uv_result$success) {
      # Only univariate succeeded
      uv_only_tbl <- uv_result$table %>%
        modify_header(label = "**Cancer Type**") %>%
        modify_table_body(~ .x %>% mutate(
          label = cancer,
          n_exposed_total = exposure_ratio
        )) %>%
        modify_header(n_exposed_total = "**N Exposed/Total**") %>%
        modify_table_body(~ .x %>% relocate(n_exposed_total, .after = label))
      
      combined_results[[cancer]] <- uv_only_tbl
      message(paste("  ⚠ Only univariate available -", exposure_ratio, "exposed"))
    }
  }
  
  # Check if we have any results
  if (length(combined_results) == 0) {
    message("\n✗ No results to export")
    return(NULL)
  }
  
  # Stack results
  stacked_table <- tbl_stack(combined_results) %>%
    modify_caption(paste0("**Cox Regression Analysis: ", drug_var, "**"))
  
  # Print table
  print(stacked_table)
  
# Export as Excel using huxtable
clean_drug_name <- gsub("_", "", tolower(drug_var))
excel_file <- here::here(output_dir, paste0("combined_", clean_drug_name, ".xlsx"))

stacked_table %>%
  as_hux_table() %>%
  huxtable::quick_xlsx(file = excel_file)

message(paste("\n✓ Saved combined Excel file:", basename(excel_file)))
  
  # Create a summary of covariates used
  covariate_summary <- tibble(
    cancer_type = names(combined_results),
    covariates_used = sapply(names(combined_results), function(cancer) {
      modified_covariates <- base_covariates
      if (cancer == "Other") {
        modified_covariates <- gsub("treatment.three", "Treatment", modified_covariates)
      }
      if (cancer == "Small cell lung cancer") {
        modified_covariates <- modified_covariates[modified_covariates != "prior.targeted_therapy"]
      }
      paste(modified_covariates, collapse = ", ")
    })
  )
  
  message("\n=== COVARIATES USED BY CANCER TYPE ===")
  print(covariate_summary)
  
  return(list(
    table = stacked_table, 
    models = model_objects,
    covariates = covariate_summary
  ))
}
```


```{r}
# RUN ANALYSES

# Create combined tables for Opioids
opioid_combined <- create_combined_tables(
  data = single_df,
  drug_var = "Opioid",
  base_covariates = c("subject.age", 
                     "prior.targeted_therapy", 
                     "subject.metastatic.disease", 
                     "stage",
                     "treatment.three")
)

# Create combined tables for Beta-Blockers
betablocker_combined <- create_combined_tables(
  data = single_df,
  drug_var = "Beta_Blockers",
  base_covariates = c("subject.age", 
                     "prior.targeted_therapy", 
                     "subject.metastatic.disease", 
                     "stage",
                     "treatment.three")
)
```


```{r}
#  Check Results
# View the tables
opioid_combined$table
betablocker_combined$table

```


```{r}
# Detailed checks

cat("\n\n=== OPIOID - SMALL CELL LUNG CANCER MODEL ===\n")
cat("Covariates in model:\n")
print(names(coef(opioid_combined$models[["Small cell lung cancer"]])))
cat("\nFull model summary:\n")
summary(opioid_combined$models[["Small cell lung cancer"]])

cat("\n\n=== OPIOID - OTHER CANCER TYPE MODEL ===\n")
cat("Covariates in model:\n")
print(names(coef(opioid_combined$models[["Other"]])))
cat("\nFull model summary:\n")
summary(opioid_combined$models[["Other"]])
```


```{r}
cat("\n\n=== BETA-BLOCKER - SMALL CELL LUNG CANCER MODEL ===\n")
cat("Covariates in model:\n")
print(names(coef(betablocker_combined$models[["Small cell lung cancer"]])))
cat("\nFull model summary:\n")
summary(betablocker_combined$models[["Small cell lung cancer"]])

cat("\n\n=== BETA-BLOCKER - OTHER CANCER TYPE MODEL ===\n")
cat("Covariates in model:\n")
print(names(coef(betablocker_combined$models[["Other"]])))
cat("\nFull model summary:\n")
summary(betablocker_combined$models[["Other"]])
```


```{r}
# check covariates used

cat("\n\n=== CHECK ALL OPIOID MODELS ===\n")
for (cancer_name in names(opioid_combined$models)) {
  cat(paste0("\n", cancer_name, ":\n"))
  cat("  Variables: ", paste(names(coef(opioid_combined$models[[cancer_name]])), collapse = ", "), "\n")
}

cat("\n\n=== CHECK ALL BETA-BLOCKER MODELS ===\n")
for (cancer_name in names(betablocker_combined$models)) {
  cat(paste0("\n", cancer_name, ":\n"))
  cat("  Variables: ", paste(names(coef(betablocker_combined$models[[cancer_name]])), collapse = ", "), "\n")
}
```



# ==============================================================================
# Beta-Blocker vs Cardiac Co-morbidity rwOS
# ==============================================================================
** Fig S1**
```{r}
# recreate cardiac condition yes/no 
cardiac_sub <- cards %>%
  select(patient.ID, cardiac_category) %>%
  rename(record_id = "patient.ID") %>%
  mutate(value = 1)

cardiac_wide <- pivot_wider(
              cardiac_sub, 
              names_from = cardiac_category, 
              values_from = value, 
              id_cols = record_id, 
              values_fill = 0, 
              values_fn = max) 

cardiac_conditions <- colnames(cardiac_wide[2:11])

cardiac_new <- cardiac_wide %>%
  mutate(
    cardiac.cond = as.integer(if_any(cardiac_conditions)))
```

```{r}
# rename
df4 <- cardiac_new 

# Merge cardiac with survival info from sub_df
sub <- sub_df %>% select(record_id, subject.dead.hospice, os, Beta_Blockers)

df5 <- left_join(sub, df4, by = c("record_id")) 

# replace NAs 

cardiac_cols <- c(
  "CORONARY ARTERY DISEASE/ISCHEMIC HEART DISEASE",
  "PERIPHERAL VASCULAR/VENOUS DISEASE (INCLUDES AORTIC ANEURYSMS)",
  "ARRYTHMIAS",
  "DYSLIPIDEMIA",
  "OTHER",
  "DVT / PULMONARY EMBOLISM",
  "VALVULAR DISEASE",
  "HEART FAILURE/CARDIOMYOPATHY",
  "CEREBROVASCULAR DISEASE",
  "EFFUSIONS",
  "cardiac.cond"
)

df5 <- df5%>%
  mutate(
    across(all_of(cardiac_cols), ~ replace_na(.x, 0))
  )

write_csv(df5, here::here("data", "processed", "multivariate_df_w_cardiac.csv"))

```



```{r}
# Contingency Table -- Beta Blocker x General Cardiac Condition

cont <- table(df5$Beta_Blockers, df5$cardiac.cond)

dimnames(cont) <- list(
  "Beta-blockers" = c("No", "Yes"),
  "Cardiac Condition" = c("No", "Yes")
)

print(cont)
```



```{r}
# Tiered, 4 KM approach
df7 <- df5 %>%
  mutate(condition = case_when(
    Beta_Blockers == 1 & cardiac.cond == 1 ~ "Beta-Blocker (Yes), Cardiac Condition (Yes)",
    Beta_Blockers == 1 & cardiac.cond == 0 ~ "Beta-Blocker (Yes), Cardiac Condition (No)",
    Beta_Blockers == 0 & cardiac.cond == 1 ~ "Beta-Blocker (No), Cardiac Condition (Yes)",
    Beta_Blockers == 0 & cardiac.cond == 0 ~ "Beta-Blocker (No), Cardiac Condition (No)")) %>%
  mutate(
    condition = factor(condition, 
                       levels = c(
                         "Beta-Blocker (Yes), Cardiac Condition (Yes)",
                         "Beta-Blocker (Yes), Cardiac Condition (No)",
                         "Beta-Blocker (No), Cardiac Condition (Yes)",
                         "Beta-Blocker (No), Cardiac Condition (No)")))

km_model <- survfit(
  Surv(os, subject.dead.hospice) ~ condition, 
  data = df7
)

```

```{r}
plot <- ggsurvplot(
  km_model,
  data = df7,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = FALSE,
  legend.title = "",
  legend.labs = levels(df7$condition),
  palette = mako(4),
  xlab = "Time",
  ylab = "Survival Probability",
  break.time.by = 6,
  surv.median.line = "hv",
  ggtheme = theme_minimal()
) 
# Apply theme and formatting to main plot
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.5, 0.2),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 18)
  ) +
  xlab("Time (months)")

# Apply theme to risk table
plot$table <- plot$table + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 18),
    plot.title = element_text(size = 20)
  ) +
  xlab("Time (months)")

print(plot)
```

```{r}
pairwise_lr <- pairwise_survdiff(
  Surv(os, subject.dead.hospice) ~ condition,
  data = df7,
  p.adjust.method = "BH"  
)

pairwise_lr

pairwise_lr$p.value

pairwise <- as.data.frame(pairwise_lr$p.value)

str(pairwise_lr)

```

```{r}
# Save the plot
output_file_png <- here::here(figures, "Fig_S1.png")
png(output_file_png, width = 11, height = 8, units = "in", res = 300)
print(plot)
dev.off()

```



# ==============================================================================
# Stage IV Opioid KM
# ==============================================================================

```{r}
# Subset to stage IV

stage4 <- sub_df %>%
  filter(stage == "iv")

```


```{r}
# Fit Kaplan-Meier model
km_model <- survfit(Surv(os, subject.dead.hospice) ~ Opioid, data = stage4)

# Define custom colors
custom_colors <- c("orange", "deepskyblue")

# Create the survival plot
plot <- ggsurvplot(
  km_model, 
  data = stage4, 
  pval = TRUE, 
  pval.size = 7,  
  risk.table.fontsize = 7, 
  conf.int = TRUE, 
  risk.table = TRUE,
  legend.labs = c("No Opioid Exposure", "Opioid Exposure"),
  palette = custom_colors,
  legend.title = "",
  title = "Stage IV"
)

# Apply theme and formatting to main plot
plot$plot <- plot$plot +
  theme_minimal() +
  theme(
    legend.position = c(0.6, 0.2),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20), 
    title = element_text(size = 22),
    plot.title = element_text(hjust = 0.5)
  ) +
  xlab("Time (months)")

# Apply theme to risk table
plot$table <- plot$table + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    plot.title = element_text(size = 20)
  ) +
  xlab("Time (months)")

print(plot)
```
```{r}
# Save the plot
output_file_png <- here::here(figures, "Reviewer_Figure_1.png")
png(output_file_png, width = 11, height = 8, units = "in", res = 300)
print(plot)
dev.off()

```

# ==============================================================================
# BB Opioid Interaction
# ==============================================================================
**Table S4**

```{r}
# Prepare data
d1_interaction <- sub_df %>%
  select(record_id, os, subject.dead.hospice, subject.age, prior.targeted_therapy,
         subject.metastatic.disease, Treatment, cancer_agreggate, stage,
         Opioid, Beta_Blockers)

# Relevel covariates
d1_interaction$Treatment <- relevel(factor(d1_interaction$Treatment), ref = "pd1-pdl1")
d1_interaction$cancer_agreggate <- relevel(factor(d1_interaction$cancer_agreggate),
                                           ref = "Non-small cell lung cancer")
d1_interaction$stage <- relevel(factor(d1_interaction$stage), ref = "iv")

# Fit model with interaction
formula_interaction <- paste0("Surv(os, subject.dead.hospice) ~ ",
                             paste(colnames(d1_interaction)[-(1:3)], collapse = " + "),
                             " + Opioid:Beta_Blockers")

fit_interaction <- coxph(as.formula(formula_interaction), data = d1_interaction)

# Summary tables
gt_table_int <- tbl_regression(fit_interaction, exponentiate = TRUE)

print(gt_table_int)

# Save as Excel files
output_dir <- here::here("output", "tables")

# Save exponentiated version (Hazard Ratios)
gt_table_int %>%
  as_hux_table() %>%
  huxtable::quick_xlsx(file = here::here(output_dir, "Table_S4_bb_interaction_model_hr.xlsx"))

```

